---
title: "Mycorrhizal Community Composition as a Barrier to Ecological Restoration"
subtitle: "Reproducible Analysis Pipeline — Hunter Mining & Snowy Hydro Case Studies"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    theme: flatly
    number_sections: true
    code_folding: hide
    fig_width: 12
    fig_height: 8
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo       = TRUE,
  warning    = FALSE,
  message    = FALSE,
  fig.align  = "center",
  fig.retina = 2,
  cache      = FALSE
)
```

# Project Setup {.tabset}

## Global Options

Toggle these flags to control what gets re-run vs loaded from cache.

```{r global-options}
# ── Reproducibility Toggles ───────────────────────────────────────────────────
# Set TRUE to rebuild processed data from raw files; FALSE loads cached RDS
RELOAD_FROM_RAW <- TRUE

# Set TRUE to re-run indicator taxa analysis; FALSE loads pre-existing CSV
RERUN_INDICATORS <- TRUE

# Set TRUE to re-compute HMSC associations & diagnostics from model objects
# (models are NEVER re-fit here — only loaded from models/ directory)
RECOMPUTE_HMSC <- TRUE

# Set TRUE to re-run CCA ordistep model fitting (slow: ~999 permutations × 4
# dataset/treatment combos). Set FALSE to load pre-fitted .RData files only.
RERUN_CCA <- FALSE
```

## Libraries

```{r libraries}
# ── Core ──────────────────────────────────────────────────────────────────────
library(tidyverse)
library(sf)
library(patchwork)

# ── Statistics ────────────────────────────────────────────────────────────────
library(lme4)
library(car)
library(broom.mixed)
library(vegan)
library(fossil)

# ── HMSC ──────────────────────────────────────────────────────────────────────
library(Hmsc)
library(coda)
library(circlize)

# ── Tables & Export ───────────────────────────────────────────────────────────
library(flextable)
library(officer)

# ── Mapping ───────────────────────────────────────────────────────────────────
library(tmap)
library(maptiles)

# ── Indicator Species ─────────────────────────────────────────────────────────
library(indicspecies)
library(permute)

# ── Plotting extras ──────────────────────────────────────────────────────────
library(ggrepel)

# Ensure dplyr::recode is available (masked by other pkgs)
recode <- dplyr::recode
select <- dplyr::select
```

## Paths

```{r paths}
# ── All paths relative to project root ────────────────────────────────────────
proj_root <- here::here()

# Raw data
its_dir <- file.path(proj_root, "rawdata/ITS")
meta_hunter <- file.path(proj_root, "rawdata/metadata/hunter")
meta_snowies <- file.path(proj_root, "rawdata/metadata/snowies")

# Pre-fitted models
models_dir <- file.path(proj_root, "models")

# Cached processed data
derived_dir <- file.path(proj_root, "deriveddata")

# Output
plot_dir <- file.path(proj_root, "output/plots")
table_dir <- file.path(proj_root, "output/tables")
indicator_csv <- file.path(table_dir, "Table_Indicator_Taxa_Combined.csv")

for (d in c(
  derived_dir, plot_dir, table_dir,
  file.path(plot_dir, "hmsc_diagnostics")
)) {
  dir.create(d, showWarnings = FALSE, recursive = TRUE)
}
```

## Publication Theme

```{r pub-theme}
# ── Shared aesthetics used across all figures ─────────────────────────────────
color_scheme <- c(Reference = "forestgreen", Rehab = "grey")

pub_theme <- theme_bw() +
  theme(
    axis.title   = element_text(size = 16, face = "bold"),
    axis.text    = element_text(size = 14, color = "black"),
    legend.title = element_text(size = 16, face = "bold"),
    legend.text  = element_text(size = 14),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    strip.text   = element_text(size = 14, face = "bold"),
    plot.title   = element_text(size = 16, face = "bold", hjust = 0.5)
  )
```

## Helper Functions

```{r helpers}
# ── Significance stars ────────────────────────────────────────────────────────
get_stars <- function(p) {
  if (is.na(p)) {
    return("")
  }
  if (p < 0.001) {
    return("***")
  }
  if (p < 0.01) {
    return("**")
  }
  if (p < 0.05) {
    return("*")
  }
  if (p < 0.1) {
    return("+")
  }
  return("ns")
}

# ── Format p-values for tables ────────────────────────────────────────────────
format_p_val <- function(p) {
  ifelse(is.na(p), "NA",
    ifelse(p < 0.0001, "< 0.0001",
      ifelse(p < 0.001, "< 0.001",
        ifelse(p <= 0.01, "< 0.01",
          as.character(round(p, 3))
        )
      )
    )
  )
}

# ── LME statistics extractor (Treatment × Patch) ─────────────────────────────
get_lme_stats <- function(data, response_var, random_effect = "(1 | Plot_ID)") {
  formula_str <- paste(response_var, "~ Treat * Patch")
  m1 <- try(lmer(as.formula(paste(formula_str, "+", random_effect)),
    data = data
  ), silent = TRUE)
  if (inherits(m1, "try-error")) {
    return(list(
      p_treat = NA, p_patch = NA, p_inter = NA,
      F_treat = NA, F_patch = NA, F_inter = NA,
      df_treat = NA, df_patch = NA, df_inter = NA,
      sig_str = "Stats Error", sig = FALSE
    ))
  }
  res <- try(tidy(Anova(m1, test = "F")), silent = TRUE)
  if (inherits(res, "try-error")) {
    return(list(
      p_treat = NA, p_patch = NA, p_inter = NA,
      F_treat = NA, F_patch = NA, F_inter = NA,
      df_treat = NA, df_patch = NA, df_inter = NA,
      sig_str = "Anova Error", sig = FALSE
    ))
  }

  extract_term <- function(term_name) {
    row <- res %>% filter(term == term_name)
    if (nrow(row) == 0) {
      return(list(p = NA, F_val = NA, df = NA))
    }
    list(
      p = row$p.value,
      F_val = round(row$statistic, 1),
      df = paste(row$df, round(row$Df.res, 0), sep = ",")
    )
  }

  t <- extract_term("Treat")
  pa <- extract_term("Patch")
  i <- extract_term("Treat:Patch")

  sig_str <- paste0(
    "Treatment ", get_stars(t$p), "\n",
    "Patch ", get_stars(pa$p), "\n",
    "Interaction ", get_stars(i$p)
  )

  list(
    p_treat = t$p, p_patch = pa$p, p_inter = i$p,
    F_treat = t$F_val, F_patch = pa$F_val, F_inter = i$F_val,
    df_treat = t$df, df_patch = pa$df, df_inter = i$df,
    sig_str = sig_str,
    sig = any(c(t$p, pa$p, i$p) < 0.05, na.rm = TRUE)
  )
}

# ── Treatment-only LME ───────────────────────────────────────────────────────
get_lme_stats_treat_only <- function(data, response_var,
                                     random_effect = "(1 | Plot_ID)") {
  formula_str <- paste(response_var, "~ Treat")
  m1 <- try(lmer(as.formula(paste(formula_str, "+", random_effect)),
    data = data
  ), silent = TRUE)
  if (inherits(m1, "try-error")) {
    return(list(
      p_treat = NA, F_treat = NA, df_treat = NA,
      sig_str = "Stats Error", sig = FALSE
    ))
  }

  res <- try(tidy(Anova(m1, test = "F")), silent = TRUE)
  if (inherits(res, "try-error")) {
    return(list(
      p_treat = NA, F_treat = NA, df_treat = NA,
      sig_str = "Anova Error", sig = FALSE
    ))
  }

  row <- res %>% filter(term == "Treat")
  if (nrow(row) == 0) {
    return(list(
      p_treat = NA, F_treat = NA, df_treat = NA,
      sig_str = "No Treat term", sig = FALSE
    ))
  }

  list(
    p_treat = row$p.value,
    F_treat = round(row$statistic, 1),
    df_treat = paste(row$df, round(row$Df.res, 0), sep = ","),
    sig_str = paste0("Treatment ", get_stars(row$p.value)),
    sig = row$p.value < 0.05
  )
}

# ── Relationship LME (y ~ x * Treat) ─────────────────────────────────────────
get_rel_stats <- function(data, y_var, x_var, random_effect) {
  formula_str <- paste(y_var, "~", x_var, "* Treat")
  m1 <- try(lmer(as.formula(paste(formula_str, "+", random_effect)),
    data = data
  ), silent = TRUE)
  if (inherits(m1, "try-error")) {
    return(list(
      p_treat = NA, p_xvar = NA, p_inter = NA,
      sig_str = "Stats Error", sig = FALSE
    ))
  }
  res <- try(tidy(Anova(m1, test = "F")), silent = TRUE)
  if (inherits(res, "try-error")) {
    return(list(
      p_treat = NA, p_xvar = NA, p_inter = NA,
      sig_str = "Anova Error", sig = FALSE
    ))
  }

  extract <- function(term_name) {
    row <- res %>% filter(term == term_name)
    if (nrow(row) == 0) {
      return(list(p = NA, F_val = NA, df = NA))
    }
    list(
      p = row$p.value, F_val = round(row$statistic, 1),
      df = paste(row$df, round(row$Df.res, 0), sep = ",")
    )
  }

  t <- extract("Treat")
  x <- extract(x_var)
  i <- extract(paste0(x_var, ":Treat"))

  sig_str <- paste0(
    "Treatment ", get_stars(t$p), "\n",
    "Covariate ", get_stars(x$p), "\n",
    "Interaction ", get_stars(i$p)
  )

  list(
    p_treat = t$p, p_xvar = x$p, p_inter = i$p,
    F_treat = t$F_val, F_xvar = x$F_val, F_inter = i$F_val,
    df_treat = t$df, df_xvar = x$df, df_inter = i$df,
    sig_str = sig_str,
    sig = any(c(t$p, x$p, i$p) < 0.05, na.rm = TRUE)
  )
}
```

# Data Loading

This section loads both the Hunter Mining and Snowy Hydro datasets.
On first run with `RELOAD_FROM_RAW = TRUE`, data is processed from raw
CSV/TSV files and cached as RDS for fast subsequent loads.

```{r load-data}
taxlev <- "genus" # taxonomic level for OTU aggregation

if (RELOAD_FROM_RAW || !file.exists(file.path(derived_dir, "hunter_dat.rds"))) {
  cat("── Loading Hunter Mining from raw data ──\n")

  # ── ITS ──
  hunter_its <- read_tsv(file.path(its_dir, "hunter_ITS_long_20210604.txt")) %>%
    mutate(Campaign = 2019) %>%
    select(Sample, Treatment, OTU, Abundance, guild, kingdom:species) %>%
    rename(Sample_ID = Sample) %>%
    mutate(
      Sample_ID = gsub("-", "", Sample_ID),
      Sample_ID = gsub("^0", "", Sample_ID)
    )

  # ── Metadata (plot + quadrat + patch) ──
  hunter_dat <- read_csv(file.path(meta_hunter, "FLATFILE-Hunter_Plot_Location_details.csv")) %>%
    mutate(across(Plot_id, as.character)) %>%
    st_as_sf(coords = c("Easting", "Northing"), crs = 32756) %>%
    st_transform(4326) %>%
    full_join(read_csv(file.path(meta_hunter, "FLATFILE-Hunter_Plot_Data.csv")) %>%
      mutate(across(Plot_id, as.character)) %>% select(-Site_id)) %>%
    full_join(read_csv(file.path(meta_hunter, "FLATFILE-Hunter_Quadrat_Data.csv"),
      na = c("", "NA", "no data")
    ) %>%
      mutate(across(Plot_id, as.character)) %>%
      rename(Establishment = Esstablishment, Survey_date = Survey_Date) %>%
      select(Plot_id, Patch, Dbh_cm:Other) %>%
      group_by(Plot_id, Patch) %>%
      summarise(
        across(`Tree sp.`, ~ paste(unique(.x), collapse = "-")),
        across(where(is.numeric), ~ mean(.x))
      ) %>%
      mutate(`Tree sp.` = ifelse(`Tree sp.` == "NA", "", `Tree sp.`))) %>%
    full_join(read_csv(file.path(meta_hunter, "FLATFILE-Hunter_Patch_Data.csv")) %>%
      mutate(across(Plot_id, as.character)) %>% select(-Treatment, -Site_id)) %>%
    mutate(
      Survey_date = lubridate::dmy(Survey_date),
      Leaf_N_P = total_nitrogen / (P_ppm / 10000),
      Treatment = ifelse(Treatment == "Rehabilitated", "Rehab", Treatment),
      Site_id = ifelse(Site_id == "Jerry's Plains", "Jerrys Plains", Site_id)
    ) %>%
    filter(!is.na(Plot_id)) %>%
    rename(
      Sample_ID = Sample_id, Treat = Treatment, Plot_ID = Plot_id, Site_ID = Site_id,
      Leaf_Mn = Mn_ppm,
      Exotic_Cover = EX_Cov, GG_cov = GG_Cov, Tree_cov = TG_Cov,
      Logs = CWD, Forb_cov = FG_Cov, Shrub_cov = SG_Cov,
      Fol_cov_short = `Foliage_Cover_<0.5`, Fol_cov_shrub = `Shrub_>0.5`,
      Fol_cov_tree = `Tree>0.5`, dist_Tree = dist_Tree_m,
      MIR_TOC = `MIR_TOC%`, P_mg_Kg = `Pmg/Kg`, EC_microS_cm = `ECmicroS/cm`,
      MIR_HUM = `MIR_HUM%`, MIR_POC = `MIR_POC%`, MIR_ROC = `MIR_ROC%`,
      MIR_TN = `MIR_TN%`,
      Total_PLFA = PLFA_Total_biomass, Fungi_Bact = `PLFA_Fungi:bacteria`,
      PHO = Enzymes_PHO, LAP = Enzymes_LAP, AB = Enzymes_AB,
      BG = Enzymes_BG, CB = Enzymes_CB, `N/AG` = Enzymes_NAG, XYL = Enzymes_XYL,
      reads.fungi = reads_fungi
    ) %>%
    select(
      Sample_ID, Treat, Plot_ID, Site_ID, PCT_Name, Patch, Establishment,
      Leaf_N_P, Leaf_Mn,
      Exotic_Cover, GG_cov, Tree_cov, Logs, Forb_cov, Shrub_cov,
      BAM_Litter, Stability_ind, Infiltration_ind, Nutrient_ind,
      Fol_cov_short, Fol_cov_shrub, Fol_cov_tree,
      MIR_TOC, P_mg_Kg, pH, EC_microS_cm, MIR_HUM, MIR_POC, MIR_ROC, MIR_TN,
      Total_PLFA, PLFA_Fungi, PLFA_Arbuscular_mycorrhizae, Fungi_Bact,
      PHO, LAP, AB, BG, CB, `N/AG`, XYL,
      reads.fungi, obs_richness_fungi_ArbuscularMycorrhizal,
      obs_richness_fungi_Ectomycorrhizal,
      relabund_fungi_ArbuscularMycorrhizal, relabund_fungi_Ectomycorrhizal
    ) %>%
    rename(plfa_amf = PLFA_Arbuscular_mycorrhizae, plfa_fungi = PLFA_Fungi) %>%
    mutate(
      plfa_amf_rel = plfa_amf / Total_PLFA,
      across(c(LAP, PHO, BAM_Litter), log10),
      relabund_fungi_Ectomycorrhizal = log10(relabund_fungi_Ectomycorrhizal + 0.0005),
      relabund_fungi_ArbuscularMycorrhizal = log10(relabund_fungi_ArbuscularMycorrhizal + 0.0005)
    ) %>%
    right_join(hunter_its %>%
      rename(taxon = matches(taxlev)) %>%
      filter(guild %in% c("Arbuscular Mycorrhizal", "Ectomycorrhizal"), !is.na(taxon)) %>%
      mutate(guild = recode(guild,
        `Arbuscular Mycorrhizal` = "AM",
        `Ectomycorrhizal` = "ECM"
      )) %>%
      unite("taxon", guild, taxon, sep = "__") %>%
      group_by(Sample_ID, taxon) %>%
      summarise(Abundance = sum(Abundance), .groups = "drop") %>%
      pivot_wider(names_from = "taxon", values_from = "Abundance", values_fill = 0)) %>%
    left_join(
      hunter_its %>%
        filter(guild %in% c("Arbuscular Mycorrhizal", "Ectomycorrhizal")) %>%
        mutate(guild = recode(guild,
          `Arbuscular Mycorrhizal` = "ArbuscularMycorrhizal",
          `Ectomycorrhizal` = "Ectomycorrhizal"
        )) %>%
        group_by(Sample_ID, guild) %>%
        summarise(chao_richness = chao1(Abundance), .groups = "drop") %>%
        pivot_wider(
          names_from = guild, values_from = chao_richness,
          names_prefix = "chao_richness_fungi_"
        ),
      by = "Sample_ID"
    ) %>%
    mutate(across(starts_with("chao_richness_fungi_"), ~ replace_na(.x, 0)))

  # Derived tables
  hunter_otus <- hunter_dat %>%
    as.data.frame() %>%
    column_to_rownames("Sample_ID") %>%
    select(starts_with("ECM__"), starts_with("AM__"))
  hunter_xy <- hunter_dat %>%
    cbind(st_coordinates(.)) %>%
    as.data.frame(row.names = .$Sample_ID) %>%
    select(Plot_ID, X, Y)
  hunter_preds <- hunter_dat %>%
    as.data.frame() %>%
    column_to_rownames("Sample_ID") %>%
    select(!c(matches(c("Plot_ID", "Site_ID", "geometry")), contains("__")))
  hunter_tax_gen <- hunter_its %>%
    rename(taxon = genus) %>%
    filter(guild %in% c("Arbuscular Mycorrhizal", "Ectomycorrhizal"), !is.na(taxon)) %>%
    mutate(guild = recode(guild,
      `Arbuscular Mycorrhizal` = "AM",
      `Ectomycorrhizal` = "ECM"
    )) %>%
    group_by(Treatment, guild, taxon) %>%
    summarise(Abundance = sum(Abundance), .groups = "drop")

  # ── Save cached ──
  saveRDS(hunter_dat, file.path(derived_dir, "hunter_dat.rds"))
  saveRDS(hunter_its, file.path(derived_dir, "hunter_its.rds"))
  saveRDS(hunter_otus, file.path(derived_dir, "hunter_otus.rds"))
  saveRDS(hunter_xy, file.path(derived_dir, "hunter_xy.rds"))
  saveRDS(hunter_preds, file.path(derived_dir, "hunter_preds.rds"))
  saveRDS(hunter_tax_gen, file.path(derived_dir, "hunter_tax_gen.rds"))
  cat("✓ Hunter data cached to deriveddata/\n")
} else {
  cat("── Loading Hunter Mining from cache ──\n")
  hunter_dat <- readRDS(file.path(derived_dir, "hunter_dat.rds"))
  hunter_its <- readRDS(file.path(derived_dir, "hunter_its.rds"))
  hunter_otus <- readRDS(file.path(derived_dir, "hunter_otus.rds"))
  hunter_xy <- readRDS(file.path(derived_dir, "hunter_xy.rds"))
  hunter_preds <- readRDS(file.path(derived_dir, "hunter_preds.rds"))
  hunter_tax_gen <- readRDS(file.path(derived_dir, "hunter_tax_gen.rds"))
}
```

```{r load-snowies}
if (RELOAD_FROM_RAW || !file.exists(file.path(derived_dir, "snowies_dat.rds"))) {
  cat("── Loading Snowy Hydro from raw data ──\n")

  snowies_its <- read_tsv(file.path(its_dir, "snowies_ITS_long_20210604.txt")) %>%
    mutate(Campaign = 2019) %>%
    select(Sample, Treatment, OTU, Abundance, guild, kingdom:species) %>%
    rename(Sample_ID = Sample)

  snowies_dat <- read_csv(file.path(meta_snowies, "SNOWIES_PLOT_DATA_2019.csv")) %>%
    mutate(across(Plot_ID, as.character)) %>%
    st_as_sf(coords = c("Easting", "Northing"), crs = 32755) %>%
    st_transform(4326) %>%
    left_join(read_csv(file.path(meta_snowies, "SnowySiteMetadata.csv")) %>%
      rename(Plot_ID = Plot, Establishment = Rehab_compl_yr) %>%
      mutate(across(Plot_ID, as.character)) %>%
      select(Plot_ID, Establishment)) %>%
    right_join(read_csv(file.path(meta_snowies, "SNOWIES_QUADRAT_DATA_2019.csv"),
      na = "N/A"
    ) %>%
      rename(
        Litter_cov = `Litter _cov`, EC_microS_cm = `EC _microS_cm`,
        Fol_cov_short = `Foliage_Cover_<0.5`,
        Fol_cov_shrub = `Fol_cov_Shrub_>0.5m`,
        Fol_cov_tree = `FC_Tree>0.5m high`
      ) %>%
      mutate(across(Plot_ID, as.character)) %>%
      select(-Elevation:-ANPP)) %>%
    select(
      Sample_ID,
      Treat = Treatment, Plot_ID, Site_ID, Subplot, Patch,
      Establishment,
      Fire_Severity_2003, ANPP,
      Leaf_N_P, Leaf_Mn,
      Exotic_Cover, Dominant_Euc, GG_cov, Tree_cov, Logs,
      Acac_ha, mean_live_DBH, CV_DBH,
      Litter_totalwt, Stability_ind, Rock, Bare_soil,
      Fol_cov_short, Fol_cov_shrub, Fol_cov_tree, dist_Tree,
      MIR_TOC, MIR_C_N, P_mg_Kg, pH, EC_microS_cm,
      Total_PLFA, Fungi_comm, arbuscular_myc_comm, Fungi_Bact,
      Abundance_inverts,
      PHO, LAP,
      reads.fungi, obs_richness_fungi_ArbuscularMycorrhizal,
      obs_richness_fungi_Ectomycorrhizal,
      relabund_fungi_ArbuscularMycorrhizal, relabund_fungi_Ectomycorrhizal,
      Forb_Cov, Shrub_cov, Euc_live_ha, Infiltration_ind, Nutrient_ind,
      Litter_cov, MIR_HUM, MIR_POC, MIR_ROC, MIR_TN, AB, BG, CB, `N/AG`, XYL
    ) %>%
    rename(plfa_amf = arbuscular_myc_comm, plfa_fungi = Fungi_comm) %>%
    mutate(
      plfa_amf_rel = plfa_amf / Total_PLFA,
      across(c(LAP, PHO, Litter_totalwt), log10),
      relabund_fungi_Ectomycorrhizal = log10(relabund_fungi_Ectomycorrhizal + 0.0005),
      relabund_fungi_ArbuscularMycorrhizal = log10(relabund_fungi_ArbuscularMycorrhizal + 0.00005)
    ) %>%
    right_join(snowies_its %>%
      rename(taxon = matches(taxlev)) %>%
      filter(guild %in% c("Arbuscular Mycorrhizal", "Ectomycorrhizal"), !is.na(taxon)) %>%
      mutate(guild = recode(guild,
        `Arbuscular Mycorrhizal` = "AM",
        `Ectomycorrhizal` = "ECM"
      )) %>%
      unite("taxon", guild, taxon, sep = "__") %>%
      group_by(Sample_ID, taxon) %>%
      summarise(Abundance = sum(Abundance), .groups = "drop") %>%
      pivot_wider(names_from = "taxon", values_from = "Abundance", values_fill = 0)) %>%
    left_join(
      snowies_its %>%
        filter(guild %in% c("Arbuscular Mycorrhizal", "Ectomycorrhizal")) %>%
        mutate(guild = recode(guild,
          `Arbuscular Mycorrhizal` = "ArbuscularMycorrhizal",
          `Ectomycorrhizal` = "Ectomycorrhizal"
        )) %>%
        group_by(Sample_ID, guild) %>%
        summarise(chao_richness = chao1(Abundance), .groups = "drop") %>%
        pivot_wider(
          names_from = guild, values_from = chao_richness,
          names_prefix = "chao_richness_fungi_"
        ),
      by = "Sample_ID"
    ) %>%
    mutate(across(starts_with("chao_richness_fungi_"), ~ replace_na(.x, 0)))

  # Derived tables
  snowies_otus <- snowies_dat %>%
    as.data.frame() %>%
    column_to_rownames("Sample_ID") %>%
    select(starts_with("ECM__"), starts_with("AM__"))
  snowies_xy <- snowies_dat %>%
    cbind(st_coordinates(.)) %>%
    as.data.frame(row.names = .$Sample_ID) %>%
    select(Plot_ID, Subplot, X, Y)
  snowies_preds <- snowies_dat %>%
    as.data.frame() %>%
    column_to_rownames("Sample_ID") %>%
    select(!c(matches(c("Plot_ID", "Site_ID", "Subplot", "geometry")), contains("__")))
  snowies_tax_gen <- snowies_its %>%
    rename(taxon = genus) %>%
    filter(guild %in% c("Arbuscular Mycorrhizal", "Ectomycorrhizal"), !is.na(taxon)) %>%
    mutate(guild = recode(guild,
      `Arbuscular Mycorrhizal` = "AM",
      `Ectomycorrhizal` = "ECM"
    )) %>%
    group_by(Treatment, guild, taxon) %>%
    summarise(Abundance = sum(Abundance), .groups = "drop")

  saveRDS(snowies_dat, file.path(derived_dir, "snowies_dat.rds"))
  saveRDS(snowies_its, file.path(derived_dir, "snowies_its.rds"))
  saveRDS(snowies_otus, file.path(derived_dir, "snowies_otus.rds"))
  saveRDS(snowies_xy, file.path(derived_dir, "snowies_xy.rds"))
  saveRDS(snowies_preds, file.path(derived_dir, "snowies_preds.rds"))
  saveRDS(snowies_tax_gen, file.path(derived_dir, "snowies_tax_gen.rds"))
  cat("✓ Snowies data cached to deriveddata/\n")
} else {
  cat("── Loading Snowy Hydro from cache ──\n")
  snowies_dat <- readRDS(file.path(derived_dir, "snowies_dat.rds"))
  snowies_its <- readRDS(file.path(derived_dir, "snowies_its.rds"))
  snowies_otus <- readRDS(file.path(derived_dir, "snowies_otus.rds"))
  snowies_xy <- readRDS(file.path(derived_dir, "snowies_xy.rds"))
  snowies_preds <- readRDS(file.path(derived_dir, "snowies_preds.rds"))
  snowies_tax_gen <- readRDS(file.path(derived_dir, "snowies_tax_gen.rds"))
}

# ── Add back-transformed EcM relative abundance for raw-scale plotting ──
hunter_dat <- hunter_dat %>%
  mutate(
    raw_relabund_fungi_Ectomycorrhizal =
      pmax(0, 10^relabund_fungi_Ectomycorrhizal - 0.0005)
  )
snowies_dat <- snowies_dat %>%
  mutate(
    raw_relabund_fungi_Ectomycorrhizal =
      pmax(0, 10^relabund_fungi_Ectomycorrhizal - 0.0005)
  )

cat("\n✓ Hunter Mining: ", nrow(hunter_dat), " samples loaded\n")
cat("✓ Snowy Hydro:   ", nrow(snowies_dat), " samples loaded\n")
```

#Study Site Maps

Interactive study site locations for both case studies using tmap view mode.
For scripts used to create figures in manuscript see R_scripts/Site_maps.R

```{r maps, fig.width=14, fig.height=7}
# ── Map helper: build an interactive tmap ────────────────────────────────────
build_figure <- function(dat, title = "") {
  pts <- dat %>%
    group_by(Plot_ID) %>%
    slice(1) %>%
    ungroup() %>%
    mutate(Treat = factor(Treat, levels = c("Reference", "Rehab")))


  tm_basemap(c("OpenStreetMap", "OpenTopoMap")) +
    tm_shape(pts) +
    tm_symbols(
      fill = "Treat",
      fill.scale = tm_scale_categorical(
        values  = color_scheme,
        labels  = c("Reference", "Restoration")
      ),
      fill.legend = tm_legend(title = paste(title, "– Treatment")),
      col = "black",
      size = .5,
      lwd = 1,
      popup.vars = c("Plot_ID", "Site_ID", "Treat")
    )
}

# ── Switch to interactive mode ────────────────────────────────────────────────
tmap_mode("view")

# ── Save static exports (plot mode) before switching ─────────────────────────
tmap_mode("plot")

hunter_static <- build_figure(hunter_dat, title = "Hunter Mining")
snowies_static <- build_figure(snowies_dat, title = "Snowy Hydro")

tmap_save(hunter_static,
  filename = file.path(plot_dir, "Figure_Map_Hunter.png"),
  width = 10, height = 8, dpi = 300
)
tmap_save(snowies_static,
  filename = file.path(plot_dir, "Figure_Map_Snowies.png"),
  width = 10, height = 8, dpi = 300
)

# ── Render interactive maps ───────────────────────────────────────────────────
# comment below if you want simple plot option
tmap_mode("view")

hunter_fig <- build_figure(hunter_dat, title = "Hunter Mining")
snowies_fig <- build_figure(snowies_dat, title = "Snowy Hydro")

hunter_static
snowies_static
```

# Soil Characteristics

Comparison of key soil abiotic properties between Reference and
Rehabilitation sites for both case studies.

```{r soil-plots, fig.width=12, fig.height=14} 
# ── Soil variable specs ───────────────────────────────────────────────────────
soil_vars <- list(
  list(var = "MIR_TOC", label = "Total Organic Carbon (%)"),
  list(var = "P_mg_Kg", label = "Colwell P (mg/kg)"),
  list(var = "Total_PLFA", label = "Total PLFA Biomass")
)

make_soil_panel <- function(data, var, ylab, title, show_x = TRUE,
                            random_eff = "(1 | Plot_ID)") {
  stats <- get_lme_stats_treat_only(data, var, random_eff)
  p <- ggplot(data, aes(x = Treat, y = !!sym(var), fill = Treat)) +
    geom_boxplot(outlier.shape = NA, linewidth = 0.8) +
    geom_jitter(width = 0.15, alpha = 0.5, size = 1.5) +
    scale_fill_manual(values = color_scheme, guide = "none") +
    labs(y = ylab, x = if (show_x) "Treatment" else "") +
    scale_x_discrete(labels = c("Reference" = "Reference", "Rehab" = "Restoration")) +
    (if (var == "P_mg_Kg") {
      scale_y_continuous(
        trans = scales::pseudo_log_trans(sigma = 25, base = 10),
        breaks = c(0, 25, 50, 100, 250, 500),
        labels = c("0", "25", "50", "100", "250", "500")
      )
    } else {
      NULL
    }) +
    # annotate("text",
    #   x = -Inf, y = Inf, label = stats$sig_str,
    #   hjust = -0.05, vjust = 1.5, size = 5, fontface = "bold" ) +
    pub_theme +
    theme(axis.text.x = if (!show_x) element_blank() else element_text(size = 16))
  p
}

# Build panels
soil_panels <- list()
for (i in seq_along(soil_vars)) {
  sv <- soil_vars[[i]]
  is_last <- i == length(soil_vars)
  p_h <- make_soil_panel(hunter_dat, sv$var, sv$label, "Hunter Mining",
    show_x = is_last
  )
  p_s <- make_soil_panel(snowies_dat, sv$var, "", "Snowy Hydro",
    show_x = is_last, random_eff = "(1 | Plot_ID / Subplot)"
  )
  soil_panels[[i]] <- p_h + p_s
}

# Column titles
title_h <- ggplot() +
  annotate("text",
    x = .5, y = .5, label = "Hunter Mining",
    size = 6, fontface = "bold"
  ) +
  theme_void()
title_s <- ggplot() +
  annotate("text",
    x = .5, y = .5, label = "Snowy Hydro",
    size = 6, fontface = "bold"
  ) +
  theme_void()
titles <- title_h + title_s

# Create tag labels for plots
tag_labels <- character(length(soil_panels) * 2)
for (i in seq_along(soil_panels)) {
  tag_labels[2 * i - 1] <- paste0(letters[i], ")") # Hunter plot gets tag
  tag_labels[2 * i] <- "" # Snowy plot gets blank
}

soil_composite <- titles / wrap_plots(soil_panels, ncol = 1) +
  plot_layout(heights = c(0.5, length(soil_vars) * 4)) +
  plot_annotation(tag_levels = list(c("", "", tag_labels))) &
  theme(plot.tag = element_text(face = "bold", size = 14))

ggsave(file.path(plot_dir, "Figure_Soil_Characteristics_Combined.png"),
  plot = soil_composite, width = 12, height = 14, dpi = 300, bg = "white"
)
print(soil_composite)
```

# Environmental PCA

Principal component analysis of environmental predictors, coloured by
treatment, with variable loadings overlaid.

```{r pca-plots, fig.width=14, fig.height=7}
# ── Variable rename map for clean axis labels ─────────────────────────────────
var_rename_map <- c(
  MIR_TOC = "TOC", P_mg_Kg = "P", pH = "pH", EC_microS_cm = "EC",
  MIR_HUM = "Humic C", MIR_POC = "POC", MIR_ROC = "ROC", MIR_TN = "TN",
  Total_PLFA = "Total PLFA", plfa_fungi = "Fungal PLFA", plfa_amf = "AMF PLFA",
  Fungi_Bact = "F:B ratio", Stability_ind = "Stability", Exotic_Cover = "Exotic cover",
  GG_cov = "Grass cover", Tree_cov = "Tree cover", Logs = "CWD",
  Fol_cov_short = "Ground cover", Fol_cov_shrub = "Shrub cover",
  Fol_cov_tree = "Tree FC", BAM_Litter = "Litter", PHO = "PHO", LAP = "LAP",
  Leaf_N_P = "Leaf N:P", Leaf_Mn = "Leaf Mn",
  relabund_fungi_Ectomycorrhizal = "EcM RA", relabund_fungi_ArbuscularMycorrhizal = "AM RA",
  Infiltration_ind = "Infiltration", Nutrient_ind = "Nutrient index",
  Litter_totalwt = "Litter wt", Rock = "Rock", Bare_soil = "Bare soil",
  dist_Tree = "Dist. tree", MIR_C_N = "C:N", Abundance_inverts = "Invertebrates",
  Fire_Severity_2003 = "Fire 2003", ANPP = "ANPP",
  Acac_ha = "Acacia density", mean_live_DBH = "Mean DBH", CV_DBH = "CV DBH"
)

run_pca_plot <- function(preds, title_text, tag_label) {
  # Mirror original: drop Establishment, fungal metrics, then NA rows
  pca_clean <- preds %>%
    select(-any_of("Establishment")) %>%
    select(!matches("relabund|richness", ignore.case = TRUE)) %>%
    drop_na()

  # Split numeric matrix and metadata (Treat stays attached)
  site_matrix <- pca_clean %>% select(where(is.numeric))
  site_meta <- pca_clean %>% select(Treat, any_of(c("Site_ID", "Patch")))

  # PCA via vegan::rda (correlation-based scaling)
  pca_res <- rda(site_matrix, scale = TRUE)

  # Variance explained
  var_exp <- summary(pca_res)$cont$importance
  axis_1 <- round(var_exp[2, 1] * 100, 1)
  axis_2 <- round(var_exp[2, 2] * 100, 1)

  # Site scores with Treat bound directly (no match() needed — rows align)
  pca_scores <- scores(pca_res, display = "sites") %>%
    as.data.frame() %>%
    bind_cols(site_meta)

  # Loadings, scaled and top 8 by magnitude
  loadings <- scores(pca_res, display = "species", choices = c(1, 2)) %>%
    as.data.frame() %>%
    rownames_to_column("Variable") %>%
    mutate(PC1 = PC1 * 1.5, PC2 = PC2 * 1.5) %>%
    mutate(loading_strength = sqrt(PC1^2 + PC2^2)) %>%
    arrange(desc(loading_strength)) %>%
    slice_head(n = 8) %>%
    mutate(Variable = dplyr::recode(Variable, !!!var_rename_map))

  ggplot() +
    geom_point(
      data = pca_scores,
      aes(x = PC1, y = PC2, color = Treat, fill = Treat),
      size = 4, alpha = 0.7, shape = 21, stroke = 0.5
    ) +
    stat_ellipse(
      data = pca_scores,
      aes(x = PC1, y = PC2, color = Treat),
      level = 0.95, linetype = 2, linewidth = 0.8
    ) +
    geom_segment(
      data = loadings,
      aes(x = 0, y = 0, xend = PC1, yend = PC2),
      arrow = arrow(length = unit(0.2, "cm"), type = "closed"),
      color = "gray30", linewidth = 0.5, alpha = 0.7
    ) +
    geom_text_repel(
      data = loadings,
      aes(x = PC1, y = PC2, label = Variable),
      size = 4, color = "gray20", fontface = "italic",
      box.padding = 0.5, max.overlaps = 20
    ) +
    scale_color_manual(
      values = color_scheme, name = "Treatment",
      labels = c("Reference" = "Reference", "Rehab" = "Restoration")
    ) +
    scale_fill_manual(
      values = color_scheme, name = "Treatment",
      labels = c("Reference" = "Reference", "Rehab" = "Restoration")
    ) +
    scale_x_continuous(expand = expansion(mult = 0.2)) +
    labs(
      x     = paste0("PC1 (", axis_1, "%)"),
      y     = paste0("PC2 (", axis_2, "%)"),
      title = title_text,
      tag   = tag_label
    ) +
    pub_theme +
    theme(
      plot.tag = element_text(size = 16, face = "bold"),
      aspect.ratio = 1
    )
}

p_pca_h <- run_pca_plot(hunter_preds, "Hunter Mining", "a)")
p_pca_s <- run_pca_plot(snowies_preds, "Snowy Hydro", "b)")

pca_composite <- p_pca_h + p_pca_s +
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom")

ggsave(file.path(plot_dir, "Figure_PCA_Combined.png"),
  plot = pca_composite, width = 14, height = 7, dpi = 300, bg = "white"
)

print(pca_composite)
```

# Mycorrhizal Community Comparisons

Box plots comparing EcM and AM richness and relative abundance

## Treatment-only Analysis

```{r myco-treatment-only, fig.height=16, fig.width=14, message=FALSE, warning=FALSE}
response_vars <- c(
  "chao_richness_fungi_Ectomycorrhizal",
  "relabund_fungi_Ectomycorrhizal",
  "chao_richness_fungi_ArbuscularMycorrhizal",
  "relabund_fungi_ArbuscularMycorrhizal"
)

var_names <- c(
  chao_richness_fungi_Ectomycorrhizal = "EcM Chao Richness",
  relabund_fungi_Ectomycorrhizal = "EcM Relative Abundance",
  chao_richness_fungi_ArbuscularMycorrhizal = "AM Chao Richness",
  relabund_fungi_ArbuscularMycorrhizal = "AM Relative Abundance (log10)"
)

combined_plots_t <- list()
stats_table_t <- data.frame()
row_counter_t <- 0

for (var in response_vars) {
  h_st <- get_lme_stats_treat_only(hunter_dat, var, "(1 | Plot_ID)")
  s_st <- get_lme_stats_treat_only(snowies_dat, var, "(1 | Plot_ID / Subplot)")

  if (h_st$sig || s_st$sig) {
    row_counter_t <- row_counter_t + 1
    is_last <- (var == response_vars[length(response_vars)])
    plot_var <- if (grepl("relabund_fungi_Ectomycorrhizal", var)) {
      paste0("raw_", var)
    } else {
      var
    }

    make_treat_box <- function(data, stats, ylab) {
      ggplot(data, aes(x = Treat, y = !!sym(plot_var), fill = Treat)) +
        geom_boxplot(outlier.shape = NA, linewidth = 1) +
        geom_jitter(width = 0.15, alpha = 0.7, size = 1.1) +
        scale_fill_manual(values = color_scheme, guide = "none") +
        scale_x_discrete(labels = c("Reference" = "Reference", "Rehab" = "Restoration")) +
        labs(x = "", y = ylab) +
        annotate("text",
          x = -Inf, y = Inf, label = stats$sig_str,
          hjust = -0.05, vjust = 1.5, size = 5, fontface = "bold"
        ) +
        pub_theme +
        theme(
          axis.text.x = if (!is_last) element_blank() else element_text(size = 14),
          axis.title.x = element_blank(),
          axis.text.y = element_text(size = 14), axis.title.y = element_text(size = 16)
        )
    }

    combined_plots_t[[var]] <-
      make_treat_box(hunter_dat, h_st, var_names[var]) +
      make_treat_box(snowies_dat, s_st, "")

    for (ds in list(list(nm = "Hunter", st = h_st), list(nm = "Snowy", st = s_st))) {
      stats_table_t <- rbind(stats_table_t, data.frame(
        Variable = var_names[var], Dataset = ds$nm,
        Treatment_df = ds$st$df_treat, Treatment_F = ds$st$F_treat,
        Treatment_p = ds$st$p_treat, stringsAsFactors = FALSE
      ))
    }
  }
}

if (length(combined_plots_t) > 0) {
  titles_t <- (ggplot() +
    annotate("text",
      x = .5, y = .5, label = "Hunter Mining",
      size = 6, fontface = "bold"
    ) +
    theme_void()) +
    (ggplot() +
      annotate("text",
        x = .5, y = .5, label = "Snowy Hydro",
        size = 6, fontface = "bold"
      ) +
      theme_void())

  n_t <- length(combined_plots_t)
  tags_t <- character(n_t * 2)
  for (i in seq_len(n_t)) {
    tags_t[2 * i - 1] <- paste0(letters[i], ")")
    tags_t[2 * i] <- ""
  }

  treat_composite <- titles_t / wrap_plots(combined_plots_t, ncol = 1) +
    plot_layout(heights = c(0.5, n_t * 4)) +
    plot_annotation(tag_levels = list(c("", "", tags_t))) &
    theme(plot.tag = element_text(face = "bold"))

  ggsave(file.path(plot_dir, "Figure_Mycorrhizal_Communities_Treatment_Only.png"),
    plot = treat_composite, width = 14, height = n_t * 4 + 2, dpi = 300, bg = "white"
  )
  print(treat_composite)
}
```

## Treatment × Patch Interaction

```{r myco-boxplots, fig.width=14, fig.height=16, warning=FALSE, message=FALSE}
combined_plots <- list()
stats_table <- data.frame()
row_counter <- 0

for (var in response_vars) {
  hunter_stats <- get_lme_stats(hunter_dat, var, "(1 | Plot_ID)")
  snowy_stats <- get_lme_stats(snowies_dat, var, "(1 | Plot_ID / Subplot)")

  if (hunter_stats$sig || snowy_stats$sig) {
    row_counter <- row_counter + 1
    is_last <- (var == "relabund_fungi_ArbuscularMycorrhizal")

    plot_var <- if (grepl("relabund_fungi_Ectomycorrhizal", var)) {
      paste0("raw_", var)
    } else {
      var
    }

    p_h <- ggplot(hunter_dat, aes(x = Patch, y = !!sym(plot_var), fill = Treat)) +
      geom_boxplot(outlier.shape = NA, linewidth = 1) +
      geom_jitter(
        position = position_jitterdodge(jitter.width = 0.1),
        alpha = 0.5, size = 1.1
      ) +
      scale_fill_manual(
        values = color_scheme,
        labels = c("Reference" = "Reference", "Rehab" = "Restoration"),
        name = "Treatment"
      ) +
      scale_x_discrete(labels = c("O" = "Open", "T" = "Beneath Trees")) +
      labs(x = "Patch", y = var_names[var]) +
      annotate("text",
        x = -Inf, y = Inf, label = hunter_stats$sig_str,
        hjust = -0.05, vjust = 1.1, size = 5, fontface = "bold", lineheight = 1
      ) +
      pub_theme +
      theme(
        legend.position = "none",
        axis.text.x = if (!is_last) element_blank() else element_text(size = 14),
        axis.title.x = if (!is_last) element_blank() else element_text(size = 16),
        axis.text.y = element_text(size = 14), axis.title.y = element_text(size = 16)
      )

    p_s <- ggplot(snowies_dat, aes(x = Patch, y = !!sym(plot_var), fill = Treat)) +
      geom_boxplot(outlier.shape = NA, linewidth = 1) +
      geom_jitter(
        position = position_jitterdodge(jitter.width = 0.1),
        alpha = 0.5, size = 1.1
      ) +
      scale_fill_manual(
        values = color_scheme,
        labels = c("Reference" = "Reference", "Rehab" = "Restoration"),
        name = "Treatment"
      ) +
      scale_x_discrete(labels = c("O" = "Open", "T" = "Beneath Trees")) +
      labs(x = "Patch", y = "") +
      annotate("text",
        x = -Inf, y = Inf, label = snowy_stats$sig_str,
        hjust = -0.05, vjust = 1.2, size = 5, fontface = "bold", lineheight = 1
      ) +
      pub_theme +
      theme(
        legend.position = "none",
        axis.text.x = if (!is_last) element_blank() else element_text(size = 14),
        axis.title.x = if (!is_last) element_blank() else element_text(size = 16),
        axis.text.y = element_text(size = 14), axis.title.y = element_text(size = 16)
      )

    combined_plots[[var]] <- p_h + p_s

    # Accumulate stats
    for (ds in list(
      list(nm = "Hunter", st = hunter_stats),
      list(nm = "Snowy", st = snowy_stats)
    )) {
      stats_table <- rbind(stats_table, data.frame(
        Variable = var_names[var], Dataset = ds$nm,
        Treatment_df = ds$st$df_treat, Treatment_F = ds$st$F_treat,
        Treatment_p = ds$st$p_treat,
        Patch_df = ds$st$df_patch, Patch_F = ds$st$F_patch,
        Patch_p = ds$st$p_patch,
        Interaction_df = ds$st$df_inter, Interaction_F = ds$st$F_inter,
        Interaction_p = ds$st$p_inter,
        stringsAsFactors = FALSE
      ))
    }
  }
}

if (length(combined_plots) > 0) {
  titles <- (ggplot() +
    annotate("text",
      x = .5, y = .5, label = "Hunter Mining",
      size = 6, fontface = "bold"
    ) +
    theme_void()) +
    (ggplot() +
      annotate("text",
        x = .5, y = .5, label = "Snowy Hydro",
        size = 6, fontface = "bold"
      ) +
      theme_void())

  n_plots <- length(combined_plots)
  tag_labels <- character(n_plots * 2)
  for (i in seq_len(n_plots)) {
    tag_labels[2 * i - 1] <- paste0(letters[i], ")")
    tag_labels[2 * i] <- ""
  }

  box_composite <- titles / wrap_plots(combined_plots, ncol = 1) +
    plot_layout(heights = c(0.5, n_plots * 4), guides = "collect") +
    plot_annotation(tag_levels = list(c("", "", tag_labels))) &
    theme(
      plot.tag = element_text(face = "bold"),
      legend.position = "bottom",
      legend.text = element_text(size = 14),
      legend.title = element_text(size = 16),
      legend.key.size = unit(1.2, "cm")
    )

  ggsave(file.path(plot_dir, "Figure_Mycorrhizal_Communities_Combined.png"),
    plot = box_composite, width = 14, height = 16, dpi = 300, bg = "white"
  )
  print(box_composite)
}
```

##  Statistics Tables for EcM and AM richness and relative abundance 

```{r myco-stats-tables}
if (nrow(stats_table) > 0) {
  st <- stats_table
  st$Treatment_p <- format_p_val(st$Treatment_p)
  st$Patch_p <- format_p_val(st$Patch_p)
  st$Interaction_p <- format_p_val(st$Interaction_p)

  ft <- flextable(st) %>%
    set_header_labels(
      Variable = "Response Variable", Dataset = "Dataset",
      Treatment_df = "df", Treatment_F = "F", Treatment_p = "p-value",
      Patch_df = "df", Patch_F = "F", Patch_p = "p-value",
      Interaction_df = "df", Interaction_F = "F", Interaction_p = "p-value"
    ) %>%
    add_header_row(
      values = c("", "", "Treatment", "Patch", "Treatment × Patch"),
      colwidths = c(1, 1, 3, 3, 3)
    ) %>%
    bold(part = "header") %>%
    autofit()

  save_as_docx(ft, path = file.path(table_dir, "Table_Mycorrhizal_Rich_RA_Stats.docx"))
}

ft
```

# Mycorrhizal Guild Relationships

Scatter plots testing the relationship between EcM and AM fungi
(richness and relative abundance) across treatments.

```{r myco-relationships, fig.width=12, fig.height=12, message=FALSE, warning=FALSE}
# Recode for these plots
hunter_dat_r <- hunter_dat %>%
  mutate(Treat = fct_recode(Treat, "Restoration" = "Rehab"))
snowies_dat_r <- snowies_dat %>%
  mutate(Treat = fct_recode(Treat, "Restoration" = "Rehab"))

relationships <- list(
  ecm_rel_vs_am_rel = list(
    y = "relabund_fungi_Ectomycorrhizal",
    x = "relabund_fungi_ArbuscularMycorrhizal",
    y_name = "EcM Relative Abundance (log10)",
    x_name = "AM Relative Abundance (log10)"
  ),
  ecm_chao_vs_am_chao = list(
    y = "chao_richness_fungi_Ectomycorrhizal",
    x = "chao_richness_fungi_ArbuscularMycorrhizal",
    y_name = "EcM Chao Richness",
    x_name = "AM Chao Richness"
  )
)

rel_plots <- list()
rel_stats <- data.frame()
rel_counter <- 0
# redefine colorscheme here
color_scheme_rehab <- c(Reference = "forestgreen", Restoration = "grey")

for (rel_name in names(relationships)) {
  rel <- relationships[[rel_name]]
  h_st <- get_rel_stats(hunter_dat_r, rel$y, rel$x, "(1 | Plot_ID)")
  s_st <- get_rel_stats(snowies_dat_r, rel$y, rel$x, "(1 | Plot_ID / Subplot)")

  if (h_st$sig || s_st$sig) {
    rel_counter <- rel_counter + 1

    make_rel_plot <- function(data, stats, ylab, tag = NULL) {
      ggplot(data, aes(x = !!sym(rel$x), y = !!sym(rel$y), color = Treat)) +
        geom_point(alpha = 0.9, size = 2) +
        geom_smooth(method = "lm", aes(fill = Treat), alpha = 0.3, linewidth = 1.2) +
        scale_color_manual(values = color_scheme_rehab, name = "Treatment") +
        scale_fill_manual(values = color_scheme_rehab, name = "Treatment") +
        labs(x = rel$x_name, y = ylab, tag = tag) +
        annotate("text",
          x = -Inf, y = Inf, label = stats$sig_str,
          hjust = -0.05, vjust = 1.1, size = 4, fontface = "bold", lineheight = 1
        ) +
        pub_theme +
        theme(
          legend.position = "bottom",
          plot.tag = element_text(size = 14, face = "bold")
        )
    }

    p_h <- make_rel_plot(
      hunter_dat_r, h_st, rel$y_name,
      paste0(letters[rel_counter], ")")
    )
    p_s <- make_rel_plot(snowies_dat_r, s_st, "")
    rel_plots[[rel_name]] <- p_h + p_s

    # Accumulate statistics for table
    for (ds in list(list(nm = "Hunter", st = h_st), list(nm = "Snowy", st = s_st))) {
      rel_stats <- rbind(rel_stats, data.frame(
        Relationship = rel$y_name,
        Covariate = rel$x_name,
        Dataset = ds$nm,
        Treatment_df = ds$st$df_treat, Treatment_F = ds$st$F_treat,
        Treatment_p = ds$st$p_treat,
        Covariate_df = ds$st$df_xvar, Covariate_F = ds$st$F_xvar,
        Covariate_p = ds$st$p_xvar,
        Interaction_df = ds$st$df_inter, Interaction_F = ds$st$F_inter,
        Interaction_p = ds$st$p_inter,
        stringsAsFactors = FALSE
      ))
    }
  }
}

if (length(rel_plots) > 0) {
  titles_r <- (ggplot() +
    annotate("text",
      x = .5, y = .5, label = "Hunter Mining",
      size = 6, fontface = "bold"
    ) +
    theme_void()) +
    (ggplot() +
      annotate("text",
        x = .5, y = .5, label = "Snowy Hydro",
        size = 6, fontface = "bold"
      ) +
      theme_void())

  rel_composite <- titles_r / wrap_plots(rel_plots, ncol = 1) +
    plot_layout(heights = c(0.5, length(rel_plots) * 6), guides = "collect") &
    theme(legend.position = "bottom")

  ggsave(file.path(plot_dir, "Figure_Mycorrhizal_Relationships_Combined.png"),
    plot = rel_composite, width = 12, height = 12, dpi = 300, bg = "white"
  )
  print(rel_composite)
}
```

## Statistics Table for Mycorrhizal Guild Relationships

```{r myco-rel-stats-table}
if (nrow(rel_stats) > 0) {
  rs <- rel_stats
  rs$Treatment_p <- format_p_val(rs$Treatment_p)
  rs$Covariate_p <- format_p_val(rs$Covariate_p)
  rs$Interaction_p <- format_p_val(rs$Interaction_p)

  ft_rel <- flextable(rs) %>%
    set_header_labels(
      Relationship   = "Relationship",
      Covariate      = "Covariate",
      Dataset        = "Dataset",
      Treatment_df   = "df",   Treatment_F   = "F",   Treatment_p   = "p-value",
      Covariate_df   = "df",   Covariate_F   = "F",   Covariate_p   = "p-value",
      Interaction_df = "df",   Interaction_F = "F",   Interaction_p = "p-value"
    ) %>%
    add_header_row(
      values    = c("", "", "", "Treatment", "Covariate", "Treatment × Covariate"),
      colwidths = c(1, 1, 1, 3, 3, 3)
    ) %>%
    align(align = "center", part = "all") %>%
    align(j = c(1, 2, 3), align = "left", part = "body") %>%
    bold(part = "header") %>%
    autofit()

  save_as_docx(ft_rel, path = file.path(
    table_dir, "Table_Mycorrhizal_Relationships_Statistics.docx"
  ))
}

ft_rel
```

# PLFA–DNA Comparison

Validation of AM fungal relative abundance estimates from ITS sequencing
against PLFA biomarker data.

```{r plfa-dna, fig.width=10, fig.height=5, message=FALSE}
# ── Hunter PLFA data ──────────────────────────────────────────────────────────
hunter_plfa <- hunter_its %>%
  left_join(
    hunter_dat %>% as.data.frame() %>%
      select(
        Sample_ID, Treat, Plot_ID, Site_ID, Patch,
        Total_PLFA, plfa_fungi, plfa_amf, reads.fungi
      ),
    by = "Sample_ID"
  ) %>%
  filter(!is.na(reads.fungi)) %>%
  mutate(Study = "Hunter Mining") %>%
  group_by(Study, Site_ID, Plot_ID, Patch, Treatment = Treat) %>%
  summarise(
    reads_fungi = sum(Abundance),
    reads_amf = sum(Abundance[guild == "Arbuscular Mycorrhizal"], na.rm = TRUE),
    plfa_total = Total_PLFA[1], plfa_fungi = plfa_fungi[1], plfa_amf = plfa_amf[1],
    .groups = "drop"
  ) %>%
  filter(!is.na(plfa_total)) %>%
  mutate(
    relabund_amf_plfa = plfa_amf / (plfa_amf + plfa_fungi),
    relabund_amf_reads_log10 = log10((reads_amf + 0.5) / (reads_fungi + 0.5))
  )

# ── Snowies PLFA data ─────────────────────────────────────────────────────────
snowies_plfa <- snowies_its %>%
  left_join(
    snowies_dat %>% as.data.frame() %>%
      select(
        Sample_ID, Treat, Plot_ID, Site_ID, Patch,
        Total_PLFA, plfa_fungi, plfa_amf, reads.fungi
      ),
    by = "Sample_ID"
  ) %>%
  filter(!is.na(reads.fungi)) %>%
  mutate(Study = "Snowy Hydro") %>%
  group_by(Study, Site_ID, Plot_ID, Patch, Treatment = Treat) %>%
  summarise(
    reads_fungi = sum(Abundance),
    reads_amf = sum(Abundance[guild == "Arbuscular Mycorrhizal"], na.rm = TRUE),
    plfa_total = Total_PLFA[1], plfa_fungi = plfa_fungi[1], plfa_amf = plfa_amf[1],
    .groups = "drop"
  ) %>%
  filter(!is.na(plfa_total)) %>%
  mutate(
    relabund_amf_plfa = plfa_amf / (plfa_amf + plfa_fungi),
    relabund_amf_reads_log10 = log10((reads_amf + 0.5) / (reads_fungi + 0.5))
  )

combined_plfa <- bind_rows(hunter_plfa, snowies_plfa) %>%
  filter(!is.na(relabund_amf_plfa), !is.na(relabund_amf_reads_log10)) %>%
  mutate(Treatment = factor(Treatment,
    levels = c("Reference", "Rehab"),
    labels = c("Reference", "Restoration")
  ))

# R² labels per group
lm_label <- function(df) {
  mod <- lm(relabund_amf_reads_log10 ~ relabund_amf_plfa, data = df)
  r2 <- summary(mod)$r.squared
  pval <- summary(mod)$coefficients[2, 4]
  sig <- case_when(
    pval < 0.001 ~ "***", pval < 0.01 ~ "**",
    pval < 0.05 ~ "*", TRUE ~ "ns"
  )
  sprintf("R² = %.2f%s", r2, sig)
}

stat_labels <- combined_plfa %>%
  group_by(Study, Treatment) %>%
  group_modify(~ tibble(label = lm_label(.x))) %>%
  ungroup() %>%
  mutate(
    x = if_else(Study == "Hunter Mining", -4.9, -4.4),
    y = case_when(
      Study == "Hunter Mining" & Treatment == "Reference" ~ 0.97,
      Study == "Hunter Mining" & Treatment == "Restoration" ~ 0.91,
      Study == "Snowy Hydro" & Treatment == "Reference" ~ 0.60,
      Study == "Snowy Hydro" & Treatment == "Restoration" ~ 0.55
    )
  )

make_plfa_plot <- function(study, tag) {
  d <- combined_plfa %>% filter(Study == study)
  sl <- stat_labels %>% filter(Study == study)
  ggplot(d, aes(
    x = relabund_amf_reads_log10, y = relabund_amf_plfa,
    colour = Treatment, fill = Treatment
  )) +
    geom_point(alpha = 0.9, size = 2.5) +
    geom_smooth(method = "lm", se = TRUE, alpha = 0.4, linewidth = 1) +
    scale_colour_manual(
      values = color_scheme_rehab,
      labels = c("Reference", "Restoration")
    ) +
    scale_fill_manual(
      values = color_scheme_rehab,
      labels = c("Reference", "Restoration")
    ) +
    labs(
      x = "AM Relative Abundance (log10 reads)",
      y = "AM Relative Abundance (PLFA)",
      title = study, tag = tag
    ) +
    geom_text(
      data = sl,
      aes(x = x, y = y, label = label, colour = Treatment),
      hjust = 0, vjust = 0, size = 3.5, fontface = "bold",
      show.legend = FALSE, inherit.aes = FALSE
    ) +
    theme_bw(base_size = 14) +
    theme(
      panel.grid.minor = element_blank(),
      plot.title = element_text(face = "bold", size = 16, hjust = 0.5),
      plot.tag = element_text(face = "bold", size = 16)
    )
}

p_plfa <- make_plfa_plot("Hunter Mining", "a)") +
  make_plfa_plot("Snowy Hydro", "b)") +
  plot_layout(guides = "collect") & theme(legend.position = "bottom")

ggsave(file.path(plot_dir, "plfa_dna_comparison_combined.png"),
  plot = p_plfa, width = 9, height = 4.5, dpi = 300, bg = "white"
)
print(p_plfa)
```

# Indicator Taxa Analysis

Indicator species analysis (IndVal) identifies taxa significantly
associated with Reference or Rehabilitation treatments.

**Note**: The pre-generated indicator taxa table is stored in
`table_dir/Table_Indicator_Taxa_Combined.csv`. Set `RERUN_INDICATORS = TRUE`
to regenerate from raw data.

```{r indicator-taxa}
run_indicator_analysis <- function(dataset_name, its_data) {
  temp <- its_data %>%
    rename(taxon = genus) %>%
    filter(guild %in% c("Arbuscular Mycorrhizal", "Ectomycorrhizal"), !is.na(taxon)) %>%
    mutate(guild = recode(guild,
      `Arbuscular Mycorrhizal` = "AM",
      `Ectomycorrhizal` = "ECM"
    )) %>%
    group_by(Sample_ID, Treatment, taxon, guild) %>%
    summarise(Abundance = sum(Abundance, na.rm = TRUE), .groups = "drop") %>%
    select(-guild) %>%
    pivot_wider(names_from = "taxon", values_from = "Abundance", values_fill = 0) %>%
    ungroup()

  res <- try(multipatt(temp %>% select(-Sample_ID, -Treatment),
    temp$Treatment,
    control = how(nperm = 999)
  ), silent = TRUE)
  if (inherits(res, "try-error")) {
    message("[indicator analysis] multipatt failed for ", dataset_name, ": ", conditionMessage(attr(res, "condition")))
    return(NULL)
  }

  species_columns <- setdiff(names(temp), c("Sample_ID", "Treatment"))

  out <- res[["sign"]] %>%
    filter(p.value <= 0.05) %>%
    rownames_to_column("taxon") %>%
    pivot_longer(cols = starts_with("s."), names_to = "group", values_to = "value") %>%
    filter(value == 1) %>%
    mutate(Indicat_Treatment = recode(gsub("^s.", "", group),
      "Rehab" = "Restoration", "REHAB" = "Restoration",
      "REFERENCE" = "Reference", "Ref" = "Reference"
    )) %>%
    select(taxon, Indicat_Treatment, stat, p.value)

  total_counts <- temp %>%
    select(all_of(species_columns)) %>%
    summarise(across(everything(), sum)) %>%
    pivot_longer(everything(), names_to = "taxon", values_to = "Total_Read_Count")

  guild_info <- its_data %>%
    rename(taxon = genus) %>%
    filter(!is.na(taxon)) %>%
    mutate(guild = recode(guild,
      `Arbuscular Mycorrhizal` = "AM",
      `Ectomycorrhizal` = "ECM"
    )) %>%
    select(taxon, guild) %>%
    distinct()

  out %>%
    left_join(total_counts, by = "taxon") %>%
    left_join(guild_info, by = "taxon") %>%
    mutate(Dataset = dataset_name) %>%
    select(Dataset, guild, Indicat_Treatment, taxon, p.value)
}

if (RERUN_INDICATORS) {
  hunter_indicators <- run_indicator_analysis("Hunter", hunter_its)
  snowy_indicators <- run_indicator_analysis("Snowy", snowies_its)

  if (!is.null(hunter_indicators) && !is.null(snowy_indicators)) {
    combined_indicators <- bind_rows(hunter_indicators, snowy_indicators) %>%
      arrange(Dataset, guild, Indicat_Treatment, taxon)
    combined_indicators$p.value <- format_p_val(combined_indicators$p.value)

    write.csv(combined_indicators, indicator_csv, row.names = FALSE)
  }
} else {
  combined_indicators <- read.csv(indicator_csv, stringsAsFactors = FALSE) %>%
    select(Dataset, guild, Indicat_Treatment, taxon, everything())
}

# Create and export flextable
common_taxa <- combined_indicators %>%
  count(taxon) %>%
  filter(n > 1) %>%
  pull(taxon)

ft_ind <- flextable(combined_indicators) %>%
  set_header_labels(
    Dataset = "Dataset", guild = "Guild", Indicat_Treatment = "Indicator of", taxon = "Genus",
    p.value = "p-value"
  ) %>%
  merge_v(j = ~ Dataset + guild + Indicat_Treatment) %>%
  theme_vanilla() %>%
  autofit() %>%
  bold(i = ~ taxon %in% common_taxa, j = "taxon", bold = TRUE)

save_as_docx(ft_ind, path = file.path(table_dir, "Table_Indicator_Taxa_Combined.docx"))
write.csv(combined_indicators, file.path(table_dir, "Table_Indicator_Taxa_Combined.csv"),
  row.names = FALSE
)
ft_ind
```

# Constrained Correspondence Analysis (CCA)

NOTE: The `ordistep()` function in vegan uses random permutation tests for
variable selection, so results are not completely reproducible across runs.
We have therefore included the original ordistep ANOVA output,
but note that re-running `ordistep` may yield slightly different variable sets.
The hardcoded variable lists below are used for the subsequent HMSC analysis to ensure full reproducibility of that modelling step.

## CCA data preparation

```{r cca-data-prep}
hunter_df <- hunter_dat %>% as.data.frame()
snowies_df <- snowies_dat %>% as.data.frame()

# Shared helper: prepare OTU + predictor matrices for a given dataset × treatment
prep_cca_data <- function(dat_df, otus, preds, treat) {
  subs <- which(dat_df$Treat == treat)
  o <- otus[subs, ]
  p <- preds[subs, ] %>%
    mutate(across(where(is.character), as.factor)) %>%
    select(where(~ sum(!is.na(.x)) >= 5)) %>%
    select(!c(
      matches(c("Plot_ID", "Site_ID", "geometry")),
      contains("__"), matches("raw_relabund|obs_richness")
    )) %>%
    as.data.frame() %>%
    select(where(~ !is.list(.x)))

  o <- o[, colSums(o) > 0]

  Y <- o[complete.cases(p), ] %>% decostand(method = "hellinger")
  Y <- Y[, colSums(Y) > 0]
  p_cc <- as.data.frame(p[complete.cases(p), ])
  list(Y = Y, p_cc = p_cc)
}

fit_cca_ordistep <- function(Y, p_cc, model_path, label) {
  # Full model
  m1_full <- cca(Y ~ ., data = p_cc)

  # Sequentially exclude variables with VIF > 10
  # vif_vals <- vif.cca(m1_full)
  # while (any(vif_vals > 10, na.rm = TRUE) || any(is.na(vif_vals))) {
  #   drop_var <- names(which.max(replace(vif_vals, is.na(vif_vals), Inf)))
  #   parent_var <- colnames(p_cc)[sapply(colnames(p_cc), function(v) startsWith(drop_var, v))]
  #   drop_col <- if (length(parent_var) > 0) parent_var[which.max(nchar(parent_var))] else drop_var
  #   message(sprintf("  [%s] Excluding '%s'  (VIF = %.1f)", label, drop_col, vif_vals[drop_var]))
  #   p_cc    <- p_cc[, setdiff(colnames(p_cc), drop_col), drop = FALSE]
  #   m1_full <- cca(Y ~ ., data = p_cc)
  #   vif_vals <- vif.cca(m1_full)
  # }

  # Bidirectional stepwise selection
  m0_full <- cca(Y ~ 1, data = p_cc)
  env_save <- new.env()
  env_save$m1.ord <- ordistep(m0_full,
    scope = formula(m1_full),
    direction = "both", permutations = how(nperm = 999), trace = FALSE
  )
  save(list = "m1.ord", file = model_path, envir = env_save)

  # Print ANOVA
  cat(sprintf(label))
  print(env_save$m1.ord$anova)

  invisible(env_save$m1.ord)
}

## CCA model fitting
# Each block below fits **one** dataset × treatment combination.
# Guarded by RERUN_CCA — set TRUE in global-options to re-run ordistep.

if (RERUN_CCA) {
  ### Hunter Mining — Reference
  d <- prep_cca_data(hunter_df, hunter_otus, hunter_preds, "Reference")
  fit_cca_ordistep(d$Y, d$p_cc,
    model_path = file.path(models_dir, "hunter/cca/ordistep_reference.RData"),
    label = "Hunter Reference"
  )
  cat("Hunter Reference predictor columns:\n")
  print(colnames(d$p_cc))

  ### Snowy Hydro — Reference
  d <- prep_cca_data(snowies_df, snowies_otus, snowies_preds, "Reference")
  fit_cca_ordistep(d$Y, d$p_cc,
    model_path = file.path(models_dir, "snowies/cca/ordistep_reference.RData"),
    label = "Snowy Reference"
  )
  cat("Snowy Reference predictor columns:\n")
  print(colnames(d$p_cc))

  ### Hunter Mining — Restoration
  d <- prep_cca_data(hunter_df, hunter_otus, hunter_preds, "Rehab")
  fit_cca_ordistep(d$Y, d$p_cc,
    model_path = file.path(models_dir, "hunter/cca/ordistep_rehab.RData"),
    label = "Hunter Restoration"
  )
  cat("Hunter Restoration predictor columns:\n")
  print(colnames(d$p_cc))

  ### Snowy Hydro — Restoration
  d <- prep_cca_data(snowies_df, snowies_otus, snowies_preds, "Rehab")
  fit_cca_ordistep(d$Y, d$p_cc,
    model_path = file.path(models_dir, "snowies/cca/ordistep_rehab.RData"),
    label = "Snowy Restoration"
  )
  cat("Snowy Restoration predictor columns:\n")
  print(colnames(d$p_cc))
} else {
  message("CCA ordistep skipped (RERUN_CCA = FALSE). Using pre-fitted .RData files.")
}

# ── Print saved ordistep ANOVA tables for reference ──────────────────────────
cca_model_specs <- list(
  list(
    path = file.path(models_dir, "hunter/cca/ordistep_reference.RData"),
    label = "Hunter Mining — Reference"
  ),
  list(
    path = file.path(models_dir, "snowies/cca/ordistep_reference.RData"),
    label = "Snowy Hydro — Reference"
  ),
  list(
    path = file.path(models_dir, "hunter/cca/ordistep_rehab.RData"),
    label = "Hunter Mining — Restoration"
  ),
  list(
    path = file.path(models_dir, "snowies/cca/ordistep_rehab.RData"),
    label = "Snowy Hydro — Restoration"
  )
)

for (spec in cca_model_specs) {
  env_loaded <- new.env()
  load(spec$path, envir = env_loaded)
  cat(sprintf("\n===== ANOVA: %s =====\n", spec$label))
  print(env_loaded$m1.ord$anova)
}

# ── Hardcoded ordistep-selected variables for HMSC ───────────────────────────
# These are the significant variables (p <= 0.05) retained by ordistep in the
# original run. They are hardcoded here so that the downstream HMSC models use
# a fixed, reviewable predictor set independent of ordistep stochasticity.

ordistep_selected_vars <- list(
  hunter_reference = c(
    "pH", "PCT_Name", "P_mg_Kg", "Stability_ind",
    "Fol_cov_shrub", "MIR_ROC", "MIR_TN", "plfa_amf",
    "EC_microS_cm", "Nutrient_ind", "Fungi_Bact"
  ),
  hunter_restoration = c(
    "Establishment", "MIR_POC", "Fol_cov_tree", "pH",
    "PCT_Name", "BAM_Litter", "LAP", "Exotic_Cover",
    "Shrub_cov", "Tree_cov"
  ),
  snowy_reference = c(
    "Exotic_Cover", "Fire_Severity_2003", "Rock",
    "Dominant_Euc", "Logs", "ANPP", "pH"
  ),
  snowy_restoration = c(
    "pH", "Leaf_N_P", "Establishment", "Tree_cov",
    "Exotic_Cover", "GG_cov", "Rock", "CV_DBH",
    "mean_live_DBH", "Logs", "Leaf_Mn", "Acac_ha",
    "dist_Tree"
  )
)

cat("\n── Hardcoded ordistep variable selections ──\n")
for (nm in names(ordistep_selected_vars)) {
  cat(sprintf("  %s: %s\n", nm, paste(ordistep_selected_vars[[nm]], collapse = ", ")))
}
```

## CCA publication plots

Build the four-panel CCA biplot figure from the saved ordistep models.

```{r cca-plots, fig.width=14, fig.height=14}
build_cca_plot <- function(dat_df, model_path, tag_label, title) {
  env_loaded <- new.env()
  load(model_path, envir = env_loaded)
  ord <- env_loaded$m1.ord

  # Eigenvalue % explained
  eig <- c(ord$CCA$eig, ord$CA$eig)
  eig_pct <- 100 * eig / sum(eig)

  # Site scores
  site_sc <- as.data.frame(scores(ord, display = "sites", choices = 1:2))
  site_sc$Sample <- rownames(site_sc)
  site_sc$Treat <- dat_df$Treat[match(site_sc$Sample, dat_df$Sample_ID)]

  # Biplot scores
  bp_sc <- as.data.frame(ord$CCA$biplot[, 1:2, drop = FALSE])
  bp_sc$var <- rownames(bp_sc)
  bp_sc$label <- ifelse(bp_sc$var %in% names(var_rename_map),
    var_rename_map[bp_sc$var], bp_sc$var
  )

  scale_f <- max(abs(site_sc[, 1:2])) * 0.6 / max(abs(bp_sc[, 1:2]))
  bp_sc[, 1:2] <- bp_sc[, 1:2] * scale_f

  ggplot() +
    geom_point(
      data = site_sc, aes(x = CCA1, y = CCA2, colour = Treat),
      size = 3, alpha = 0.7
    ) +
    geom_segment(
      data = bp_sc, aes(x = 0, y = 0, xend = CCA1, yend = CCA2),
      arrow = arrow(length = unit(0.2, "cm")),
      colour = "black", alpha = 0.6, linewidth = 0.6
    ) +
    geom_text(
      data = bp_sc, aes(x = CCA1, y = CCA2, label = label),
      size = 3.5, fontface = "bold", vjust = -0.5
    ) +
    scale_colour_manual(
      values = color_scheme,
      name = if (title %in% c("Hunter Mining — Restoration", "Snowy Hydro — Restoration")) NULL else "Treatment",
      labels = c("Reference" = "Reference", "Rehab" = "Restoration")
    ) +
    labs(
      x = paste0("CCA1 (", round(eig_pct[1], 1), "%)"),
      y = paste0("CCA2 (", round(eig_pct[2], 1), "%)"),
      title = title, tag = tag_label
    ) +
    pub_theme +
    theme(plot.tag = element_text(size = 16, face = "bold"), aspect.ratio = 1)
}

cca_plots <- list(
  build_cca_plot(
    hunter_df,
    file.path(models_dir, "hunter/cca/ordistep_reference.RData"),
    "a)", "Hunter Mining — Reference"
  ),
  build_cca_plot(
    snowies_df,
    file.path(models_dir, "snowies/cca/ordistep_reference.RData"),
    "b)", "Snowy Hydro — Reference"
  ),
  build_cca_plot(
    hunter_df,
    file.path(models_dir, "hunter/cca/ordistep_rehab.RData"),
    "c)", "Hunter Mining — Restoration"
  ),
  build_cca_plot(
    snowies_df,
    file.path(models_dir, "snowies/cca/ordistep_rehab.RData"),
    "d)", "Snowy Hydro — Restoration"
  )
)

cca_plots <- Filter(Negate(is.null), cca_plots)
if (length(cca_plots) > 0) {
  cca_composite <- wrap_plots(cca_plots, ncol = 2) +
    plot_layout(guides = "collect") & theme(legend.position = "bottom")

  ggsave(file.path(plot_dir, "publication_cca_combined.png"),
    plot = cca_composite, width = 14, height = 14, dpi = 300, bg = "white"
  )
  print(cca_composite)
}
```
# HMSC Model Fitting (Optional — Reproducibility)

The HMSC models are computationally intensive and were originally fit on a
remote server. Pre-fitted model objects are loaded from the `models/` directory
throughout this document. To regenerate any model from scratch, uncomment the
corresponding `source()` call below. Each script preps the data, fits both PA
and abundance models via MCMC, evaluates convergence, and saves outputs to the
paths expected by this document.

**Note:** Running these scripts will take several hours per model.

```{r hmsc-run-models, eval=FALSE}
# ── Uncomment to re-fit individual HMSC models ────────────────────────────────
# Each script preps data, fits PA + abundance models, computes convergence
# diagnostics, and saves outputs to models/<study>/<treatment>/

# Hunter Mining — Reference sites
source(file.path(proj_root, "R_scripts/hmsc/dataPrep_hmsc_hunter_reference.R"))

# Hunter Mining — Rehab (Restoration) sites
source(file.path(proj_root, "R_scripts/hmsc/dataPrep_hmsc_hunter_rehab.R"))

# Snowy Hydro — Reference sites
source(file.path(proj_root, "R_scripts/hmsc/dataPrep_hmsc_snowies_reference.R"))

# Snowy Hydro — Rehab (Restoration) sites
source(file.path(proj_root, "R_scripts/hmsc/dataPrep_hmsc_snowies_rehab.R"))
```

# HMSC Model Diagnostics

Diagnostic plots for HMSC presence-absence and abundance models,
including PSRF convergence, model fit metrics, and effective sample sizes.

```{r hmsc-diagnostics, fig.width=10, fig.height=10, results='hide'}
# ── PA Diagnostic Function ────────────────────────────────────────────────────
plot_diagnostics_PA <- function(mpost, psrf_beta, mf_expl, label) {
  out_file <- file.path(
    plot_dir, "hmsc_diagnostics",
    paste0("Diagnostics_PA_", label, ".pdf")
  )
  pdf(out_file, width = 10, height = 10)
  par(mfrow = c(2, 2), mar = c(5, 5, 4, 2))

  try({
    hist(psrf_beta[, 1],
      main = paste("PSRF Beta — PA", label),
      xlab = "PSRF", col = "steelblue", border = "white", breaks = 30
    )
    abline(v = 1.1, col = "red", lty = 2, lwd = 2)
    legend("topright", legend = c(
      paste("Mean =", round(mean(psrf_beta[, 1]), 4)),
      paste(
        "Range:", round(min(psrf_beta[, 1]), 4), "–",
        round(max(psrf_beta[, 1]), 4)
      )
    ), bty = "n", cex = 0.9)
  })

  if (!is.null(mf_expl$TjurR2)) {
    barplot(sort(mf_expl$TjurR2, decreasing = TRUE),
      main = paste("Tjur R²", label), ylab = "Tjur R²",
      col = "steelblue", border = NA, las = 2, cex.names = 0.4
    )
    abline(h = mean(mf_expl$TjurR2, na.rm = TRUE), col = "red", lty = 2)
  } else {
    plot.new()
    text(0.5, 0.5, "Tjur R² N/A")
  }

  if (!is.null(mf_expl$AUC)) {
    barplot(sort(mf_expl$AUC, decreasing = TRUE),
      main = paste("AUC", label), ylab = "AUC",
      col = "darkorange", border = NA, las = 2, cex.names = 0.4
    )
    abline(h = 0.5, col = "grey50", lty = 3)
    abline(h = mean(mf_expl$AUC, na.rm = TRUE), col = "red", lty = 2)
  } else {
    plot.new()
    text(0.5, 0.5, "AUC N/A")
  }

  try({
    ess <- effectiveSize(mpost$Beta)
    hist(ess,
      main = paste("ESS Beta", label), xlab = "ESS",
      col = "mediumpurple", border = "white", breaks = 30
    )
    legend("topright",
      legend = paste("Median =", round(median(ess), 1)),
      bty = "n", cex = 0.9
    )
  })

  dev.off()
}

# ── Abundance Diagnostic Function ────────────────────────────────────────────
plot_diagnostics_AB <- function(mpost, psrf_beta, mf_expl, mf_pred, label) {
  out_file <- file.path(
    plot_dir, "hmsc_diagnostics",
    paste0("Diagnostics_Abundance_", label, ".pdf")
  )
  pdf(out_file, width = 10, height = 10)
  par(mfrow = c(2, 2), mar = c(5, 5, 4, 2))

  try({
    hist(psrf_beta[, 1],
      main = paste("PSRF Beta — Abundance", label),
      xlab = "PSRF", col = "steelblue", border = "white", breaks = 30
    )
    abline(v = 1.1, col = "red", lty = 2, lwd = 2)
  })

  if (!is.null(mf_expl$R2)) {
    if (!is.null(mf_pred) && !is.null(mf_pred$R2)) {
      plot(mf_expl$R2, mf_pred$R2,
        xlim = c(-0.1, 1), ylim = c(-0.1, 1),
        xlab = "Explanatory R²", ylab = "Predictive R²",
        main = paste("R²", label), pch = 21, bg = "forestgreen"
      )
      abline(0, 1, col = "red", lty = 2)
    } else {
      barplot(sort(mf_expl$R2, decreasing = TRUE),
        main = paste("R² (expl)", label), col = "forestgreen", border = NA
      )
    }
  } else {
    plot.new()
    text(0.5, 0.5, "R² N/A")
  }

  if (!is.null(mf_expl$RMSE)) {
    if (!is.null(mf_pred) && !is.null(mf_pred$RMSE)) {
      rng <- range(c(mf_expl$RMSE, mf_pred$RMSE), na.rm = TRUE)
      plot(mf_expl$RMSE, mf_pred$RMSE,
        xlim = rng, ylim = rng,
        xlab = "Explanatory RMSE", ylab = "Predictive RMSE",
        main = paste("RMSE", label), pch = 21, bg = "grey50"
      )
      abline(0, 1, col = "red", lty = 2)
    }
  } else {
    plot.new()
    text(0.5, 0.5, "RMSE N/A")
  }

  try({
    ess <- effectiveSize(mpost$Beta)
    hist(ess,
      main = paste("ESS Beta", label), xlab = "ESS",
      col = "mediumpurple", border = "white", breaks = 30
    )
  })

  dev.off()
}

# ── Run diagnostics for all 4 model sets ─────────────────────────────────────
diag_specs <- list(
  list(path = "snowies/hmsc_reference/mcmc_convergence.RData", label = "Snowy_Reference"),
  list(path = "snowies/hmsc_rehab/mcmc_convergence.RData", label = "Snowy_Rehab"),
  list(path = "hunter/hmsc_reference/mcmc_convergence.RData", label = "Hunter_Reference"),
  list(path = "hunter/hmsc_rehab/mcmc_convergence.RData", label = "Hunter_Rehab")
)

for (spec in diag_specs) {
  env <- new.env()
  load(file.path(models_dir, spec$path), envir = env)
  plot_diagnostics_PA(env$mpost1, env$psrf.beta1, env$MF1, spec$label)
  plot_diagnostics_AB(env$mpost2, env$psrf.beta2, env$MF2, env$MF_pred, spec$label)
}

cat("All HMSC diagnostic plots saved to output/plots/hmsc_diagnostics/\n")
```

# HMSC Model Evaluation Table

Summary table of model fit metrics (AUC, Tjur R², R², RMSE) and MCMC
convergence diagnostics across all model × treatment combinations.

```{r hmsc-eval-table}
get_vp <- function(hM) {
  vp <- computeVariancePartitioning(hM)
  rowMeans(vp$vals, na.rm = TRUE)
}

get_model_fit <- function(hM) {
  predY <- computePredictedValues(hM)
  mf <- evaluateModelFit(hM, predY = predY)
  tibble(
    AUC = if (!is.null(mf$AUC)) mean(mf$AUC, na.rm = TRUE) else NA_real_,
    TjurR2 = if (!is.null(mf$TjurR2)) mean(mf$TjurR2, na.rm = TRUE) else NA_real_,
    Mean_R2 = if (!is.null(mf$R2)) mean(mf$R2, na.rm = TRUE) else NA_real_,
    RMSE = if (!is.null(mf$RMSE)) mean(mf$RMSE, na.rm = TRUE) else NA_real_
  )
}

get_psrf <- function(hM) {
  mpost <- convertToCodaObject(hM)
  psrf_beta <- gelman.diag(mpost$Beta, multivariate = FALSE)$psrf
  list(
    mean = sprintf("%.4f", mean(psrf_beta[, 1])),
    range = sprintf("%.4f–%.4f", min(psrf_beta[, 1]), max(psrf_beta[, 1]))
  )
}

build_eval_table <- function(models, label) {
  fit_PA <- get_model_fit(models[[1]])
  fit_AB <- get_model_fit(models[[2]])
  psrf_PA <- get_psrf(models[[1]])
  psrf_AB <- get_psrf(models[[2]])

  tibble(
    Metric = c("AUC", "Tjur R²", "Mean R²", "RMSE", "Mean PSRF (β)", "PSRF range (β)"),
    PA = c(
      round(fit_PA$AUC, 3), round(fit_PA$TjurR2, 3), "--", "--",
      psrf_PA$mean, psrf_PA$range
    ),
    AB = c(
      "--", "--", round(fit_AB$Mean_R2, 3), round(fit_AB$RMSE, 3),
      psrf_AB$mean, psrf_AB$range
    )
  ) %>% setNames(c("Metric", paste0(label, "_PA"), paste0(label, "_Abundance")))
}

# Load models and build tables
eval_specs <- list(
  list(path = "hunter/hmsc_reference/mcmc_output.RData", label = "Hunter_Ref"),
  list(path = "hunter/hmsc_rehab/mcmc_output.RData", label = "Hunter_Rehab"),
  list(path = "snowies/hmsc_reference/mcmc_output.RData", label = "Snowy_Ref"),
  list(path = "snowies/hmsc_rehab/mcmc_output.RData", label = "Snowy_Rehab")
)

eval_tables <- list()
for (spec in eval_specs) {
  env <- new.env()
  load(file.path(models_dir, spec$path), envir = env)
  eval_tables[[spec$label]] <- build_eval_table(env$models, spec$label)
}

# Combine into one table
combined_eval <- eval_tables[[1]]
for (i in 2:length(eval_tables)) {
  combined_eval <- full_join(combined_eval, eval_tables[[i]], by = "Metric")
}
combined_eval[is.na(combined_eval)] <- "--"

# Build flextable
ft_eval <- flextable(combined_eval) %>%
  set_header_df(mapping = tibble(
    col_keys = colnames(combined_eval),
    line1 = c("", rep("Hunter Mining", 4), rep("Snowy Hydro", 4)),
    line2 = c(
      "", rep("Reference", 2), rep("Restoration", 2),
      rep("Reference", 2), rep("Restoration", 2)
    ),
    line3 = c("Metric", rep(c("PA", "Abundance"), 4))
  )) %>%
  theme_booktabs() %>%
  merge_h(part = "header") %>%
  merge_v(part = "header") %>%
  align(align = "center", part = "header") %>%
  align(j = 1, align = "left", part = "all") %>%
  bold(part = "header") %>%
  autofit()

doc_eval <- read_docx() %>%
  body_add_par("HMSC Model Evaluation Summary", style = "heading 1") %>%
  body_add_flextable(ft_eval)
print(doc_eval, target = file.path(table_dir, "Table_HMSC_Model_Evaluation.docx"))

ft_eval
```

# HMSC Chord Diagrams

Chord diagrams showing significant co-occurrence associations between
mycorrhizal fungal taxa from HMSC models.

```{r hmsc-chord, fig.width=10, fig.height=10, results='hide'}
# ── Helper functions ──────────────────────────────────────────────────────────
tidymat <- function(x) {
  rownames(x) <- gsub("[A-Z]+\\_\\_", "", rownames(x))
  colnames(x) <- gsub("[A-Z]+\\_\\_", "", colnames(x))
  x
}

compute_associations <- function(models, model_idx, dataset = c("snowy", "hunter")) {
  dataset <- match.arg(dataset)
  m <- models[[model_idx]]
  OmegaCor <- computeAssociations(m, start = 1)
  rl_names <- names(m$ranLevels)
  target_level <- if (dataset == "snowy") "Subplot" else "Plot_ID"
  ranlev <- which(rl_names == target_level)[1]
  supportLevel <- 0.6
  ((OmegaCor[[ranlev]]$support > supportLevel) +
    (OmegaCor[[ranlev]]$support < (1 - supportLevel)) > 0) * OmegaCor[[ranlev]]$mean
}

draw_chord <- function(toPlot, title_text, targets, label = NULL) {
  adjm <- toPlot
  diag(adjm) <- 0
  all_taxa <- rownames(toPlot)
  keep <- unique(unlist(lapply(targets, function(t) grep(t, all_taxa, value = TRUE))))
  if (length(keep) == 0 || !any(adjm[keep, , drop = FALSE] != 0)) {
    plot.new()
    mtext(title_text, side = 3, line = 0.5, cex = 0.9, font = 2)
    text(0.5, 0.5, "(No significant associations)")
    return(invisible(NULL))
  }
  adjm <- adjm[keep, , drop = FALSE]
  other <- colnames(adjm)[!colnames(adjm) %in% keep]
  cols <- c(
    rep("#5f0854", length(keep)),
    ifelse(grepl("ECM__", other), "#a06f99", "#b68317")
  )
  names(cols) <- c(keep, other)
  names(cols) <- gsub("[A-Z]+\\_\\_", "", names(cols))
  adjm <- tidymat(adjm)
  cols <- cols[names(cols) %in% c(rownames(adjm), colnames(adjm))]
  circos.par(canvas.xlim = c(-1.0, 1.0), canvas.ylim = c(-1.0, 1.0))
  chordDiagram(adjm,
    col = function(x) ifelse(x < 0, "red", "blue"),
    grid.col = cols, annotationTrack = "grid",
    preAllocateTracks = list(track.height = max(strwidth(unlist(dimnames(adjm)))) * 0.6)
  )
  mtext(title_text, side = 3, line = 1.5, cex = 0.9, font = 2)
  if (!is.null(label)) mtext(label, side = 3, adj = 0, line = 1.5, cex = 0.9, font = 2)
  circos.track(track.index = 1, track.margin = c(0.06, 0), panel.fun = function(x, y) {
    circos.text(CELL_META$xcenter, CELL_META$ylim[1], CELL_META$sector.index,
      facing = "clockwise", niceFacing = TRUE, adj = c(0, .2), cex = .65
    )
  }, bg.border = NA)

  circos.clear()
}

draw_chord_am_only <- function(toPlot, title_text, targets, label = NULL) {
  adjm <- toPlot
  diag(adjm) <- 0
  keep <- unique(unlist(lapply(targets, function(t) grep(t, rownames(toPlot), value = TRUE))))
  am_cols <- grep("^AM__", colnames(adjm), value = TRUE)
  if (length(keep) == 0 || length(am_cols) == 0 || !any(adjm[keep, am_cols, drop = FALSE] != 0)) {
    plot.new()
    mtext(title_text, side = 3, line = 0.5, cex = 0.9, font = 2)
    text(0.5, 0.5, "(No significant EcM-AM associations)")
    return(invisible(NULL))
  }
  adjm <- adjm[keep, am_cols, drop = FALSE]
  row_clean <- gsub("[A-Z]+\\_\\_", "", rownames(adjm))
  col_clean <- gsub("[A-Z]+\\_\\_", "", colnames(adjm))
  cols <- c(
    setNames(rep("#5f0854", length(row_clean)), row_clean),
    setNames(rep("#b68317", length(col_clean)), col_clean)
  )
  rownames(adjm) <- row_clean
  colnames(adjm) <- col_clean
  cols <- cols[unique(names(cols))]
  circos.par(canvas.xlim = c(-1.0, 1.0), canvas.ylim = c(-1.0, 1.0))
  chordDiagram(adjm,
    col = function(x) ifelse(x < 0, "red", "blue"),
    grid.col = cols, annotationTrack = "grid",
    preAllocateTracks = list(track.height = max(strwidth(unlist(dimnames(adjm)))) * 0.6)
  )
  mtext(title_text, side = 3, line = 1.5, cex = 0.9, font = 2)
  if (!is.null(label)) mtext(label, side = 3, adj = 0, line = 1.5, cex = 0.9, font = 2)
  circos.track(track.index = 1, track.margin = c(0.06, 0), panel.fun = function(x, y) {
    circos.text(CELL_META$xcenter, CELL_META$ylim[1], CELL_META$sector.index,
      facing = "clockwise", niceFacing = TRUE, adj = c(0, .2), cex = .65
    )
  }, bg.border = NA)
  circos.clear()
}

has_significant_am <- function(toPlot, targets) {
  adjm <- toPlot
  diag(adjm) <- 0
  keep <- unique(unlist(lapply(targets, function(t) grep(t, rownames(toPlot), value = TRUE))))
  am_cols <- grep("^AM__", colnames(adjm), value = TRUE)
  if (length(keep) == 0 || length(am_cols) == 0) {
    return(FALSE)
  }
  any(adjm[keep, am_cols, drop = FALSE] != 0)
}

has_significant_chord <- function(toPlot, targets) {
  adjm <- toPlot
  diag(adjm) <- 0
  keep <- unique(unlist(lapply(targets, function(t) grep(t, rownames(toPlot), value = TRUE))))
  if (length(keep) == 0) {
    return(FALSE)
  }
  any(adjm[keep, , drop = FALSE] != 0)
}

# ── Load indicator taxa for shared targets ────────────────────────────────────
indicator_df <- read.csv(indicator_csv, stringsAsFactors = FALSE)
indicator_df$Indicat_Treatment_Norm <- toupper(indicator_df$Indicat_Treatment)
indicator_df$Indicat_Treatment_Norm[indicator_df$Indicat_Treatment_Norm == "REF"] <- "REFERENCE"

shared_ref_targets <- intersect(
  indicator_df %>% filter(Dataset == "Hunter", Indicat_Treatment_Norm == "REFERENCE") %>%
    pull(taxon) %>% unique(),
  indicator_df %>% filter(Dataset == "Snowy", Indicat_Treatment_Norm == "REFERENCE") %>%
    pull(taxon) %>% unique()
)

# ── Compute associations ─────────────────────────────────────────────────────
assoc_specs <- list(
  list(path = "snowies/hmsc_reference/mcmc_output.RData", ds = "snowy", label = "snowy_ref"),
  list(path = "snowies/hmsc_rehab/mcmc_output.RData", ds = "snowy", label = "snowy_rehab"),
  list(path = "hunter/hmsc_reference/mcmc_output.RData", ds = "hunter", label = "hunter_ref"),
  list(path = "hunter/hmsc_rehab/mcmc_output.RData", ds = "hunter", label = "hunter_rehab")
)

assoc <- list()
for (spec in assoc_specs) {
  env <- new.env()
  load(file.path(models_dir, spec$path), envir = env)
  assoc[[paste0(spec$label, "_pa")]] <- compute_associations(env$models, 1, spec$ds)
  assoc[[paste0(spec$label, "_ab")]] <- compute_associations(env$models, 2, spec$ds)
}

# ── Draw combined chord figure ────────────────────────────────────────────────
chord_pairs <- list(
  list(
    hunter = list(data = assoc$hunter_ref_pa, title = "Hunter — Ref (PA)"),
    snowy = list(data = assoc$snowy_ref_pa, title = "Snowy — Ref (PA)")
  ),
  list(
    hunter = list(data = assoc$hunter_ref_ab, title = "Hunter — Ref (Abund)"),
    snowy = list(data = assoc$snowy_ref_ab, title = "Snowy — Ref (Abund)")
  ),
  list(
    hunter = list(data = assoc$hunter_rehab_pa, title = "Hunter — Restoration (PA)"),
    snowy = list(data = assoc$snowy_rehab_pa, title = "Snowy — Restoration (PA)")
  ),
  list(
    hunter = list(data = assoc$hunter_rehab_ab, title = "Hunter — Restoration (Abund)"),
    snowy = list(data = assoc$snowy_rehab_ab, title = "Snowy — Restoration (Abund)")
  )
)

# Keep only pairs where at least one dataset has ANY significant association
# (positive or negative, across all fungal guilds — not restricted to EcM-AM).
valid_pairs <- Filter(function(pr) {
  has_significant_chord(pr$hunter$data, shared_ref_targets) ||
    has_significant_chord(pr$snowy$data, shared_ref_targets)
}, chord_pairs)

if (length(valid_pairs) > 0) {
  output_chord <- file.path(plot_dir, "hmsc_chord_combined.pdf")
  fig_h <- length(valid_pairs) * 3.5
  pdf(output_chord, width = 10, height = fig_h)
  par(mfrow = c(length(valid_pairs), 2), mar = c(1, 1, 2, 1), oma = c(1, 0, 1, 0))
  idx <- 1
  for (pr in valid_pairs) {
    # Draw all significant associations for each site (Hunter then Snowy)
    draw_chord(
      pr$hunter$data, pr$hunter$title, shared_ref_targets,
      paste0(letters[idx], ")")
    )
    idx <- idx + 1
    draw_chord(
      pr$snowy$data, pr$snowy$title, shared_ref_targets,
      paste0(letters[idx], ")")
    )
    idx <- idx + 1
  }
  invisible(dev.off())
}

# ── Save to file (AM-only associations) ──────────────────────────────────────
output_chord_rehab <- file.path(plot_dir, "hmsc_chord_rehab_pa.png")
png(output_chord_rehab, width = 10, height = 5, units = "in", res = 300)
par(mfrow = c(1, 2), mar = c(1, 1, 3, 1), oma = c(0, 0, 0, 0))
# Save Restoration (PA) chord diagrams showing only significant AM fungal associations
for (info in list(
  list(data = assoc$hunter_rehab_pa, title = "Hunter Mining — Restoration (PA)", label = "a)"),
  list(data = assoc$snowy_rehab_pa, title = "Snowy Hydro — Restoration (PA)", label = "b)")
)) {
  draw_chord_am_only(info$data, info$title, shared_ref_targets, info$label)
}
invisible(dev.off())

# ── Inline display ────────────────────────────────────────────────────────────
par(mfrow = c(1, 2), mar = c(1, 1, 3, 1), oma = c(0, 0, 0, 0))
# Inline display: Restoration (PA) chord diagrams with only significant AM fungal associations
for (info in list(
  list(data = assoc$hunter_rehab_pa, title = "Hunter Mining — Restoration (PA)", label = "a)"),
  list(data = assoc$snowy_rehab_pa, title = "Snowy Hydro — Restoration (PA)", label = "b)")
)) {
  draw_chord_am_only(info$data, info$title, shared_ref_targets, info$label)
}
par(mfrow = c(1, 1))
```

# HMSC Interactive Chord Diagrams

Interactive HTML chord diagrams showing positive and negative fungal co-occurrence
associations side-by-side, separated by sign. Blue chords = positive; Red chords = negative.
Dark nodes = focal EcM indicator taxa; Medium = other EcM; Light = AM fungi.

```{r hmsc-chord-interactive, results='asis'}
# ── Packages needed for interactive output ────────────────────────────────────
# remotes::install_github("mattflor/chorddiag")   # if not on CRAN
if (!requireNamespace("chorddiag", quietly = TRUE) ||
  !requireNamespace("htmlwidgets", quietly = TRUE) ||
  !requireNamespace("base64enc", quietly = TRUE)) {
  knitr::knit_exit()
}

output_dir <- plot_dir # maps to output/plots/

# ── Helper: filter association matrix to keep only positive or negative values ─
filter_by_sign <- function(mat, sign_type) {
  filtered <- mat
  if (sign_type == "positive") {
    filtered[filtered < 0] <- 0
  } else if (sign_type == "negative") {
    filtered[filtered > 0] <- 0
  }
  return(filtered)
}

# ── Build one chorddiag widget ─────────────────────────────────────────────────
create_interactive_chord <- function(toPlot, targets, sign_type = "all") {
  adjm <- toPlot
  diag(adjm) <- 0

  adjm <- filter_by_sign(adjm, sign_type)

  all_taxa <- rownames(toPlot)
  keep <- unique(unlist(lapply(targets, function(t) grep(t, all_taxa, value = TRUE))))
  if (length(keep) == 0) {
    return(NULL)
  }

  adjm <- adjm[keep, , drop = FALSE]
  if (!any(adjm != 0)) {
    return(NULL)
  }

  adjm <- adjm[, colSums(abs(adjm)) > 0, drop = FALSE]

  if (sign_type == "positive") {
    ecm_dark <- "#041331"
    ecm_light <- "#2a5090"
    am_col <- "#c7e0f7"
  } else if (sign_type == "negative") {
    ecm_dark <- "#4e0909"
    ecm_light <- "#9a2020"
    am_col <- "#f3d5d5"
  } else {
    ecm_dark <- "#5f0854"
    ecm_light <- "#a06f99"
    am_col <- "#b68317"
  }

  other <- colnames(adjm)[!colnames(adjm) %in% keep]
  row_cols <- setNames(rep(ecm_dark, length(keep)), keep)
  col_cols <- setNames(ifelse(grepl("ECM__", other), ecm_light, am_col), other)
  color_map <- c(row_cols, col_cols)

  names(color_map) <- gsub("[A-Z]+\\_\\_", "", names(color_map))
  color_map <- color_map[!duplicated(names(color_map))]

  adjm <- tidymat(adjm)

  all_names <- unique(c(rownames(adjm), colnames(adjm)))
  square_adjm <- matrix(0,
    nrow = length(all_names), ncol = length(all_names),
    dimnames = list(all_names, all_names)
  )
  square_adjm[rownames(adjm), colnames(adjm)] <- adjm

  display_adjm <- abs(square_adjm)
  display_adjm <- pmax(display_adjm, t(display_adjm))

  has_links <- (rowSums(display_adjm) + colSums(display_adjm)) > 0
  display_adjm <- display_adjm[has_links, has_links, drop = FALSE]
  all_names <- rownames(display_adjm)
  if (length(all_names) == 0) {
    return(NULL)
  }

  group_colors <- color_map[all_names]
  group_colors[is.na(group_colors)] <- "#999999"

  chord_col <- if (sign_type == "positive") "#3366cc" else if (sign_type == "negative") "#cc3333" else "#888888"

  chorddiag::chorddiag(
    display_adjm,
    groupColors = unname(group_colors),
    chordedgeColor = chord_col,
    groupnamePadding = 30,
    groupnameFontsize = 11,
    showTicks = FALSE,
    margin = 120,
    tooltipGroupConnector = " \u2194 ",
    tooltipUnit = "",
    precision = 3,
    fadeLevel = 0.3
  )
}

# ── Build self-contained HTML page (no pandoc required) ───────────────────────
html_output <- file.path(output_dir, "publication_hmsc_chord_interactive.html")

html_parts <- c(
  "<!DOCTYPE html>",
  "<html><head>",
  '<meta charset="utf-8">',
  "<title>HMSC Chord Diagrams - Fungal Associations</title>",
  "<style>body{max-width:1400px;margin:0 auto;padding:20px;font-family:Arial,sans-serif;}",
  ".pair-row{display:flex;flex-wrap:wrap;gap:20px;justify-content:center;margin-bottom:30px;}",
  ".pair-col{flex:1;min-width:400px;}",
  "h3{margin-top:30px;} hr{margin:20px 0;}</style>",
  "</head><body>",
  "<h1>HMSC Chord Diagrams - Fungal Associations</h1>",
  "<p>Each panel shows positive (blue, left) and negative (red, right) associations. ",
  "Dark shade = EcM indicator taxa; Medium shade = Other ECM; Light shade = AM fungi.</p>"
)

# Collect widgets for inline display in the knitted HTML
widget_list <- list()

if (length(valid_pairs) > 0) {
  label_idx2 <- 1
  for (pr in valid_pairs) {
    for (s in list(pr$hunter, pr$snowy)) {
      # Use shared_ref_targets (the targets computed in the previous chunk)
      w_pos <- create_interactive_chord(s$data, shared_ref_targets, sign_type = "positive")
      w_neg <- create_interactive_chord(s$data, shared_ref_targets, sign_type = "negative")

      if (is.null(w_pos) && is.null(w_neg)) {
        label_idx2 <- label_idx2 + 1
        next
      }

      # ── Accumulate for inline knit display ───────────────────────────────────
      section_heading <- htmltools::h3(paste0(letters[label_idx2], ") ", s$title))
      inline_row <- htmltools::div(
        style = "display:flex;flex-wrap:wrap;gap:20px;justify-content:center;margin-bottom:30px;",
        htmltools::div(
          style = "flex:1;min-width:400px;",
          htmltools::h4(
            style = "text-align:center;color:#1a3a6b;",
            "Positive Associations"
          ),
          if (!is.null(w_pos)) {
            w_pos
          } else {
            htmltools::p(
              style = "color:#999;font-style:italic;text-align:center;",
              "(No positive associations)"
            )
          }
        ),
        htmltools::div(
          style = "flex:1;min-width:400px;",
          htmltools::h4(
            style = "text-align:center;color:#8b1a1a;",
            "Negative Associations"
          ),
          if (!is.null(w_neg)) {
            w_neg
          } else {
            htmltools::p(
              style = "color:#999;font-style:italic;text-align:center;",
              "(No negative associations)"
            )
          }
        )
      )
      widget_list <- c(widget_list, list(section_heading, inline_row, htmltools::hr()))

      html_parts <- c(html_parts, paste0(
        "<h3>", letters[label_idx2], ") ", s$title, "</h3>"
      ))
      html_parts <- c(html_parts, '<div class="pair-row">')

      # ── Positive column ──
      html_parts <- c(html_parts, '<div class="pair-col">')
      html_parts <- c(html_parts, '<h4 style="text-align:center;color:#1a3a6b;">Positive Associations</h4>')
      if (!is.null(w_pos)) {
        pos_tmp <- tempfile(fileext = ".html")
        htmlwidgets::saveWidget(w_pos, file = pos_tmp, selfcontained = TRUE)
        html_parts <- c(html_parts, paste0(
          '<iframe src="data:text/html;base64,',
          base64enc::base64encode(charToRaw(paste(readLines(pos_tmp, warn = FALSE), collapse = "\n"))),
          '" style="width:100%;height:520px;border:none;"></iframe>'
        ))
        unlink(pos_tmp)
      } else {
        html_parts <- c(html_parts, '<p style="color:#999;font-style:italic;text-align:center;">(No positive associations)</p>')
      }
      html_parts <- c(html_parts, "</div>")

      # ── Negative column ──
      html_parts <- c(html_parts, '<div class="pair-col">')
      html_parts <- c(html_parts, '<h4 style="text-align:center;color:#8b1a1a;">Negative Associations</h4>')
      if (!is.null(w_neg)) {
        neg_tmp <- tempfile(fileext = ".html")
        htmlwidgets::saveWidget(w_neg, file = neg_tmp, selfcontained = TRUE)
        html_parts <- c(html_parts, paste0(
          '<iframe src="data:text/html;base64,',
          base64enc::base64encode(charToRaw(paste(readLines(neg_tmp, warn = FALSE), collapse = "\n"))),
          '" style="width:100%;height:520px;border:none;"></iframe>'
        ))
        unlink(neg_tmp)
      } else {
        html_parts <- c(html_parts, '<p style="color:#999;font-style:italic;text-align:center;">(No negative associations)</p>')
      }
      html_parts <- c(html_parts, "</div>")

      html_parts <- c(html_parts, "</div><hr>")
      label_idx2 <- label_idx2 + 1
    }
  }
} else {
  cat("  No significant associations found for interactive chord diagram.\n")
}

html_parts <- c(html_parts, "</body></html>")

writeLines(paste(html_parts, collapse = "\n"), con = html_output)

if (file.exists(html_output)) {
  cat("✓ Interactive HTML saved (self-contained):", html_output, "\n")
} else {
  cat("[ERROR] Failed to write interactive HTML to:", html_output, "\n")
}

htmltools::tagList(widget_list)
```

#Crosshair plots showing  relationship between EcM indicator genera RA and environmental variables

Reference site means ± SD are shown as crosshairs, with rehabilitation site data as
scatter points with regression lines.

```{r genera-recovery, fig.width=14, fig.height=10}
# ── Target genera from indicator taxa ─────────────────────────────────────────
indicator_df$Indicat_Treatment_Norm <- toupper(indicator_df$Indicat_Treatment)
indicator_df$Indicat_Treatment_Norm[indicator_df$Indicat_Treatment_Norm == "REF"] <- "REFERENCE"

hunter_ecm_ref <- indicator_df %>%
  filter(Dataset == "Hunter", guild == "ECM", Indicat_Treatment_Norm == "REFERENCE") %>%
  pull(taxon) %>%
  unique()

snowy_ecm_ref <- indicator_df %>%
  filter(Dataset == "Snowy", guild == "ECM", Indicat_Treatment_Norm == "REFERENCE") %>%
  pull(taxon) %>%
  unique()

ecm_ref_indicators <- intersect(hunter_ecm_ref, snowy_ecm_ref)

# Criterion: EcM genus must (1) be a significant indicator of reference sites
# in both case studies AND (2) show a significant negative residual association
# with AM fungi in the HMSC rehab model (posterior support > 0.60).
extract_negative_ecm_am <- function(assoc_matrix, targets = NULL) {
  if (is.null(assoc_matrix)) {
    return(data.frame(EcM = character(), AM = character(), stringsAsFactors = FALSE))
  }
  ecm_rows <- grep("^ECM__", rownames(assoc_matrix), value = TRUE)
  am_cols <- grep("^AM__", colnames(assoc_matrix), value = TRUE)
  if (!is.null(targets) && length(targets) > 0) {
    ecm_rows <- unique(unlist(lapply(targets, function(t) grep(t, ecm_rows, value = TRUE))))
  }
  if (length(ecm_rows) == 0 || length(am_cols) == 0) {
    return(data.frame(EcM = character(), AM = character(), stringsAsFactors = FALSE))
  }
  sub_mat <- assoc_matrix[ecm_rows, am_cols, drop = FALSE]
  neg_pairs <- which(sub_mat < 0, arr.ind = TRUE)
  if (nrow(neg_pairs) == 0) {
    return(data.frame(EcM = character(), AM = character(), stringsAsFactors = FALSE))
  }
  data.frame(
    EcM = gsub("ECM__", "", rownames(sub_mat)[neg_pairs[, 1]]),
    AM = gsub("AM__", "", colnames(sub_mat)[neg_pairs[, 2]]),
    stringsAsFactors = FALSE
  )
}

snowy_neg_pa <- extract_negative_ecm_am(assoc$snowy_rehab_pa, ecm_ref_indicators)
snowy_neg_ab <- extract_negative_ecm_am(assoc$snowy_rehab_ab, ecm_ref_indicators)
hunter_neg_pa <- extract_negative_ecm_am(assoc$hunter_rehab_pa, ecm_ref_indicators)
hunter_neg_ab <- extract_negative_ecm_am(assoc$hunter_rehab_ab, ecm_ref_indicators)

all_neg_pairs <- rbind(
  if (nrow(snowy_neg_pa) > 0) cbind(snowy_neg_pa, Dataset = "Snowy", Model = "PA"),
  if (nrow(snowy_neg_ab) > 0) cbind(snowy_neg_ab, Dataset = "Snowy", Model = "Abundance"),
  if (nrow(hunter_neg_pa) > 0) cbind(hunter_neg_pa, Dataset = "Hunter", Model = "PA"),
  if (nrow(hunter_neg_ab) > 0) cbind(hunter_neg_ab, Dataset = "Hunter", Model = "Abundance")
)

snowy_ecm_targets <- unique(c(snowy_neg_pa$EcM, snowy_neg_ab$EcM))
hunter_ecm_targets <- unique(c(hunter_neg_pa$EcM, hunter_neg_ab$EcM))

# ── Site recovery variable lists ──────────────────────────────────────────────
site_vars_common <- list(
  Tree_cov = list(
    col = "Tree_cov",
    label = "Native Tree Cover (%)"
  ),
  pH = list(
    col = "pH",
    label = "Soil pH"
  ),
  P_mg_Kg = list(
    col = "P_mg_Kg",
    label = "Available Phosphorus (mg/kg)"
  )
)
selected_var_labels <- c(
  "Native Tree Cover (%)",
  "Soil pH",
  "Available Phosphorus (mg/kg)"
)


get_lm_stats <- function(data, y_col, x_col) {
  rehab_data <- data %>% filter(Treat == "Rehab")
  y_vals <- rehab_data[[y_col]]
  x_vals <- rehab_data[[x_col]]
  complete <- !is.na(y_vals) & !is.na(x_vals)
  if (sum(complete) < 5) {
    return(list(
      p = NA, r2 = NA, r = NA, f = NA, df1 = NA, df2 = NA,
      n = sum(complete), sig_str = "n < 5"
    ))
  }
  m <- try(lm(as.formula(paste(y_col, "~", x_col)), data = rehab_data), silent = TRUE)
  if (inherits(m, "try-error")) {
    return(list(
      p = NA, r2 = NA, r = NA, f = NA, df1 = NA, df2 = NA,
      n = nrow(rehab_data), sig_str = "Model error"
    ))
  }
  s <- summary(m)
  p_val <- tryCatch(s$coefficients[2, 4], error = function(e) NA)
  r2 <- s$r.squared
  fstat <- s$fstatistic
  list(
    p = p_val, r2 = round(r2, 3),
    f = if (!is.null(fstat)) round(fstat[1], 2) else NA,
    df1 = if (!is.null(fstat)) fstat[2] else NA,
    df2 = if (!is.null(fstat)) fstat[3] else NA,
    n = sum(complete),
    sig_str = paste0(
      "R\u00B2 = ", round(r2, 2), ", p = ",
      formatC(p_val, format = "g", digits = 2)
    )
  )
}

make_crosshair_plot <- function(data, genus_name, genus_col, x_col, x_label,
                                guild_prefix = "ECM", colour = "forestgreen",
                                add_lm = TRUE) {
  y_col <- paste0(".y_relabund_", genus_name)
  data <- data %>% mutate(!!y_col := log10((!!sym(genus_col) + 0.5) / reads.fungi))
  ref_s <- data %>%
    filter(Treat == "Reference") %>%
    summarise(
      y_mn = mean(!!sym(y_col), na.rm = TRUE),
      y_sd = sd(!!sym(y_col), na.rm = TRUE),
      x_mn = mean(!!sym(x_col), na.rm = TRUE),
      x_sd = sd(!!sym(x_col), na.rm = TRUE)
    )
  if (is.na(ref_s$y_mn) || is.na(ref_s$x_mn)) {
    return(NULL)
  }
  rehab <- data %>% filter(Treat == "Rehab")
  if (nrow(rehab) < 3) {
    return(NULL)
  }
  p <- ggplot(rehab, aes(x = !!sym(x_col), y = !!sym(y_col))) +
    geom_point(colour = colour, alpha = 0.6, size = 2)
  if (add_lm) {
    p <- p + geom_smooth(
      method = "lm", colour = colour,
      fill = alpha(colour, 0.3), linewidth = 0.8, se = TRUE
    )
  }
  p + geom_pointrange(
    data = ref_s,
    aes(x = x_mn, y = y_mn, xmin = x_mn - x_sd, xmax = x_mn + x_sd),
    inherit.aes = FALSE, colour = "forestgreen", linewidth = 1.2, size = 1.2
  ) +
    geom_pointrange(
      data = ref_s,
      aes(x = x_mn, y = y_mn, ymin = y_mn - y_sd, ymax = y_mn + y_sd),
      inherit.aes = FALSE, colour = "forestgreen", linewidth = 1.2, size = 1.2
    ) +
    labs(x = x_label, y = paste0("RA of ", genus_name, " (log10)")) +
    theme_classic() +
    theme(
      axis.title = element_text(size = 14, face = "bold"),
      axis.text = element_text(size = 12),
      plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
      plot.margin = margin(4, 8, 4, 4)
    )
}

run_genera_analysis <- function(target_genera_hunter, target_genera_snowy,
                                guild_prefix, guild_label,
                                hunter_data, snowy_data,
                                site_vars_list_hunter, site_vars_list_snowy,
                                colour = "grey40") {
  all_genera <- unique(c(target_genera_hunter, target_genera_snowy))
  all_plots <- list()
  stats_results <- data.frame()

  for (genus in all_genera) {
    genus_col <- paste0(guild_prefix, "__", genus)
    has_hunter <- genus_col %in% names(hunter_data)
    has_snowy <- genus_col %in% names(snowy_data)
    if (!has_hunter && !has_snowy) next

    genus_plots_hunter <- list()
    genus_plots_snowy <- list()

    make_site_plots <- function(data, site_vars, dataset_label) {
      plots <- list()
      new_stats <- data.frame()
      for (var_name in names(site_vars)) {
        var_info <- site_vars[[var_name]]
        if (!var_info$col %in% names(data)) next
        p <- make_crosshair_plot(
          data, genus, genus_col, var_info$col, var_info$label,
          guild_prefix, colour
        )
        if (!is.null(p)) {
          y_tmp <- paste0(".y_relabund_", genus)
          d_tmp <- data %>% mutate(!!y_tmp := log10((!!sym(genus_col) + 0.5) / reads.fungi))
          stats <- get_lm_stats(d_tmp, y_tmp, var_info$col)
          new_stats <- rbind(new_stats, data.frame(
            Genus = genus, Guild = guild_label, Site_Variable = var_info$label,
            Dataset = dataset_label, n = stats$n, R_squared = stats$r2,
            F_statistic = stats$f, df1 = stats$df1, df2 = stats$df2,
            p_value = stats$p, Significance = get_stars(stats$p),
            stringsAsFactors = FALSE
          ))
          p <- p + annotate("text",
            x = -Inf, y = Inf, label = stats$sig_str,
            hjust = -0.05, vjust = 1.3, size = 4, colour = "grey30"
          )
          plots[[var_name]] <- p
        }
      }
      list(plots = plots, stats = new_stats)
    }

    res_hunter <- if (has_hunter) make_site_plots(hunter_data, site_vars_list_hunter, "Hunter") else list(plots = list(), stats = data.frame())
    res_snowy <- if (has_snowy) make_site_plots(snowy_data, site_vars_list_snowy, "Snowy") else list(plots = list(), stats = data.frame())

    stats_results <- rbind(stats_results, res_hunter$stats, res_snowy$stats)
    genus_plots_hunter <- res_hunter$plots
    genus_plots_snowy <- res_snowy$plots

    all_plots[[genus]] <- list(hunter = genus_plots_hunter, snowy = genus_plots_snowy)
  }
  list(plots = all_plots, stats = stats_results)
}

# ── Run analyses ───────────────────────────────────────────────────────────────
ecm_results <- run_genera_analysis(
  target_genera_hunter = hunter_ecm_targets, target_genera_snowy = snowy_ecm_targets,
  guild_prefix = "ECM", guild_label = "EcM",
  hunter_data = hunter_dat, snowy_data = snowies_dat,
  site_vars_list_hunter = site_vars_common, site_vars_list_snowy = site_vars_common,
  colour = "grey40"
)

# p value threshold
threshold <- 0.1

create_genus_figure <- function(genus_name, genus_plots, guild_label, colour,
                                stats_df = NULL, sig_threshold = 1,
                                site_vars_lookup = site_vars_common) {
  hunter_plots <- genus_plots$hunter
  snowy_plots <- genus_plots$snowy
  if (!is.null(stats_df) && nrow(stats_df) > 0) {
    key_to_label <- sapply(site_vars_lookup, function(x) x$label)
    all_var_keys <- unique(c(names(hunter_plots), names(snowy_plots)))
    keep_vars <- c()
    for (var_key in all_var_keys) {
      var_label <- key_to_label[[var_key]]
      sig_h <- stats_df %>% filter(Genus == genus_name, Dataset == "Hunter", Site_Variable == var_label)
      sig_s <- stats_df %>% filter(Genus == genus_name, Dataset == "Snowy", Site_Variable == var_label)
      hunter_sig <- nrow(sig_h) > 0 && !is.na(sig_h$p_value[1]) && sig_h$p_value[1] < sig_threshold
      snowy_sig <- nrow(sig_s) > 0 && !is.na(sig_s$p_value[1]) && sig_s$p_value[1] < sig_threshold
      if (hunter_sig || snowy_sig) keep_vars <- c(keep_vars, var_key)
    }
    hunter_plots <- hunter_plots[intersect(names(hunter_plots), keep_vars)]
    snowy_plots <- snowy_plots[intersect(names(snowy_plots), keep_vars)]
  }
  if (length(hunter_plots) == 0 && length(snowy_plots) == 0) {
    return(NULL)
  }
  all_var_names <- unique(c(names(hunter_plots), names(snowy_plots)))
  n_rows <- length(all_var_names)
  title_hunter <- ggplot() +
    annotate("text", x = 0.5, y = 0.5, label = "Hunter Mining", size = 5, fontface = "bold") +
    theme_void()
  title_snowy <- ggplot() +
    annotate("text", x = 0.5, y = 0.5, label = "Snowy Hydro", size = 5, fontface = "bold") +
    theme_void()
  plot_list <- list(title_hunter, title_snowy)
  for (var in all_var_names) {
    p_h <- if (var %in% names(hunter_plots)) {
      hunter_plots[[var]]
    } else {
      ggplot() +
        theme_void() +
        annotate("text", x = 0.5, y = 0.5, label = "(ns)", size = 4, colour = "grey50")
    }
    p_s <- if (var %in% names(snowy_plots)) {
      snowy_plots[[var]]
    } else {
      ggplot() +
        theme_void() +
        annotate("text", x = 0.5, y = 0.5, label = "(ns)", size = 4, colour = "grey50")
    }
    plot_list <- c(plot_list, list(p_h, p_s))
  }
  wrap_plots(plot_list, ncol = 2, nrow = n_rows + 1, heights = c(0.15, rep(1, n_rows))) +
    plot_annotation(theme = theme(plot.title = element_text(size = 14, face = "bold", hjust = 0.5)))
}

output_dir <- file.path(proj_root, "output/plots")

ecm_genus_figures <- list()
for (genus in names(ecm_results$plots)) {
  fig <- create_genus_figure(genus, ecm_results$plots[[genus]], "EcM", "grey40",
    stats_df = ecm_results$stats, sig_threshold = threshold,
    site_vars_lookup = site_vars_common
  )
  if (!is.null(fig)) {
    ecm_genus_figures[[genus]] <- fig
  }
}

# ── EcM master figure (all significant panels, p < 0.1) ──────────────────────
build_study_column <- function(all_results_list, study_name, stats_df, site_vars_list) {
  study_key <- ifelse(study_name == "Hunter", "hunter", "snowy")
  plots <- list()
  for (genus in names(all_results_list)) {
    genus_study_plots <- all_results_list[[genus]][[study_key]]
    if (is.null(genus_study_plots) || length(genus_study_plots) == 0) next
    for (var_name in names(genus_study_plots)) {
      var_label <- site_vars_list[[var_name]]$label
      sig_row <- stats_df %>% filter(Genus == genus, Dataset == study_name, Site_Variable == var_label)
      if (nrow(sig_row) > 0 && !is.na(sig_row$p_value) && sig_row$p_value < threshold) {
        plots <- c(plots, list(genus_study_plots[[var_name]]))
      }
    }
  }
  plots
}

ecm_hunter_col <- build_study_column(ecm_results$plots, "Hunter", ecm_results$stats, site_vars_common)
ecm_snowy_col <- build_study_column(ecm_results$plots, "Snowy", ecm_results$stats, site_vars_common)
max_plots <- max(length(ecm_hunter_col), length(ecm_snowy_col))

title_hunter <- ggplot() +
  annotate("text", x = 0.5, y = 0.5, label = "Hunter Mining", size = 5, fontface = "bold") +
  theme_void()
title_snowy <- ggplot() +
  annotate("text", x = 0.5, y = 0.5, label = "Snowy Hydro", size = 5, fontface = "bold") +
  theme_void()

ecm_master_list <- list(title_hunter, title_snowy)
for (i in seq_len(max_plots)) {
  p_h <- if (i <= length(ecm_hunter_col)) {
    ecm_hunter_col[[i]]
  } else {
    ggplot() +
      theme_void()
  }
  p_s <- if (i <= length(ecm_snowy_col)) {
    ecm_snowy_col[[i]]
  } else {
    ggplot() +
      theme_void()
  }
  ecm_master_list <- c(ecm_master_list, list(p_h, p_s))
}

if (length(ecm_master_list) > 2) {
  tags <- c(paste0(letters, ")"), paste0("a", letters, ")"))
  plot_idx <- 1
  for (i in 3:length(ecm_master_list)) {
    row_i <- floor((i - 3) / 2) + 1
    col_i <- ((i - 3) %% 2) + 1
    is_real_hunter <- col_i == 1 && row_i <= length(ecm_hunter_col)
    is_real_snowy <- col_i == 2 && row_i <= length(ecm_snowy_col)
    if (is_real_hunter || is_real_snowy) {
      ecm_master_list[[i]] <- ecm_master_list[[i]] + labs(tag = tags[plot_idx])
      plot_idx <- plot_idx + 1
    }
  }
  ecm_master <- wrap_plots(ecm_master_list,
    ncol = 2,
    heights = c(0.12, rep(1, max_plots))
  ) &
    theme(plot.tag = element_text(face = "bold"), plot.tag.margin = margin(0, 8, 0, 0))
  ggsave(file.path(output_dir, "Figure_EcM_Master_SiteRecovery.png"),
    plot = ecm_master, width = 14, height = min(4 * max_plots + 2, 300),
    dpi = 300, bg = "white", limitsize = FALSE
  )
}

# ── Statistics table ───────────────────────────────────────────────────────────
all_stats <- rbind(ecm_results$stats) %>% #
  arrange(Guild, Dataset, Genus, Site_Variable)

if (nrow(all_stats) > 0) {
  all_stats$p_formatted <- format_p_val(all_stats$p_value)
  all_stats$df <- ifelse(is.na(all_stats$df1) | is.na(all_stats$df2), NA,
    paste0(all_stats$df1, ", ", all_stats$df2)
  )
  display_cols <- c(
    "Genus", "Guild", "Site_Variable", "Dataset", "n",
    "R_squared", "F_statistic", "df", "p_formatted"
  )
  ft <- flextable(all_stats %>% select(-p_value, -df1, -df2), col_keys = display_cols) %>%
    set_header_labels(
      Genus = "Genus", Guild = "Guild", Site_Variable = "Site Variable",
      Dataset = "Dataset", n = "n (Rehab)", R_squared = "R\u00B2",
      F_statistic = "F", df = "df", p_formatted = "p-value"
    ) %>%
    align(align = "center", part = "all") %>%
    align(j = c(1, 2, 3, 4), align = "left", part = "body") %>%
    bold(part = "header") %>%
    italic(j = "Genus", part = "body") %>%
    autofit() %>%
    bold(i = ~ Significance %in% c("+", "*", "**", "***"), j = "p_formatted")
  tables_dir <- file.path(proj_root, "output/tables")
  dir.create(tables_dir, showWarnings = FALSE, recursive = TRUE)
  save_as_docx(ft, path = file.path(tables_dir, "Table_Genera_SiteRecovery_Statistics.docx"))
  sig_results <- all_stats %>% filter(Significance %in% c("*", "**", "***"))
}

# ── Display figures ────────────────────────────────────────────────────────────
ecm_master
# Display stats
ft
```


# Session Info

```{r session-info}
sessionInfo()
```
